### 2022-09-06 SticklesEDA.Rmd
Commiting Data and initial EDA

---
title: "Stickleback EDA"
author: "Akhil Ghosh"
date: "2022-09-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Loading Libraries
```{r}
library(tidyverse)
library(ggplot2)
library(readr)
library(dplyr)
library(DT)

```

```{r}
stickle <- read_csv('stickleback_fossil_vs_extant.csv')
summary(stickle)
#seems every column outside of ID has missing data to some extent
#Highest amount of NA's is cle.sc column
#least amount of NAs in mds column
```

```{r}
datatable(stickle)

#Should I split up univID?
#Maybe, maybe not
```
```{r}
stickles <- datatable(stickle)
grepl("M",stickle$univID, fixed=TRUE)
stickle$univID
stick <- stickle %>% 
  mutate(gender = ifelse(grepl("M",univID,fixed=TRUE),"M","F"))

table(stick$gender)
#Only 130 Males?????
#sheesh

```