### 2022-09-21 SticklesEDA.Rmd
lots of stickle changes

---
title: "Stickleback EDA"
author: "Akhil Ghosh"
date: "2022-09-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Loading Libraries
```{r}
library(tidyverse)
library(DT)

```

```{r}
stickle <- read_csv('stickleback_fossil_vs_extant.csv')
summary(stickle)
#seems every column outside of ID has missing data to some extent
#Highest amount of NA's is cle.sc column
#least amount of NAs in mds column
```

```{r}
datatable(stickle)

#Should I split up univID?
#Maybe, maybe not
```
```{r}
stickles <- datatable(stickle)
grepl("M",stickle$univID, fixed=TRUE)
stickle$univID

stick <- stickle %>% 
  mutate(gender = case_when(
    grepl(".M.",univID,fixed=TRUE) ~ "Male",
    grepl(".F.",univID,fixed=TRUE) ~ "Female",
    grepl("Fos",univID,fixed=TRUE) ~ NA_character_
    )
  )
           # ifelse(grepl("M",univID,fixed=TRUE),"M","F"))

table(stick$gender)
#Only 284 extant fish

summary(stick)

summary(as.factor(stick$gender))

```
```{r}
install.packages("mice")
library(mice)

md.pattern(stickle)

```

```{r}
# devtools::install_github("agnesdeng/mixgb")

library(mixgb)

stick_clean <- stick %>% 
  select(.,-univID) %>% 
  mutate(ips = as.factor(ips))

stick_clean <- data_clean(stick_clean)
stick_clean
```
```{r}
# imputed_data <- mixgb(data = stick_clean, m = 5)
# imputed_data

cv.results <- mixgb_cv(data = stick_clean, verbose = FALSE)
cv.results$response
cv.results$best.nrounds

cv.results2 <- mixgb_cv(data = stick_clean, nfold = 10, nrounds = 100,
    early_stopping_rounds = 1, response = "gender", verbose = FALSE)
cv.results2$best.nrounds

imputed_data2 <- mixgb(data = stick_clean, m = 5, nrounds = 10)
imputed_data3 <- mixgb(data = stick_clean, m = 5, nrounds = 30)
imputed_data3

```

```{r}

#Here I try to impute every variable in the data set and then use the imputed dataset to predict gender variable. Through this I can cross validate and avoid data cleaning issues.

#Just a prelim thing I wanted to see the accuracy
#For this, I will be focusing on the extant dataset.

summary(stick_clean)

non_gender <- stick_clean %>% 
  select(-gender)

cv.results_ <- mixgb_cv(data = non_gender, verbose=FALSE)
cv.results_$best.nrounds
cv.results_$response

imputed_data_no_gender <- mixgb(data = no, m = 5, nrounds = 100)
imputed_data_ng <- mixgb(data = non_gender, m=5, nrounds = 100)

imputed_data_ng

```

```{r}
best_rounds <- c()
for(i in 1:50){
cv.results_ <- mixgb_cv(data = non_gender, verbose=FALSE,nrounds=200)
best_rounds <-append(best_rounds, cv.results_$best.nrounds)
if(best_rounds[i]==max(best_rounds)){
  maxresponse = cv.results_$response
  maxRounds <- cv.results_$best.nrounds
}

}

max(best_rounds)
maxresponse

cv.results_ <- mixgb_cv(data = non_gender,response="stl", verbose=FALSE,nrounds=200)
cv.results_$best.nrounds

parallel::detectCores()
n.cores <- parallel::detectCores() - 1

my.cluster <- parallel::makeCluster(
  n.cores, 
  type = "FORK"
  )
doParallel::registerDoParallel(cl = my.cluster)
foreach::getDoParRegistered()
foreach::getDoParWorkers()

best_rounds<-
x <- foreach(
  i = 1:100, 
  .combine = 'c'
) %dopar% {
  cv.results_ <- mixgb_cv(data = non_gender, verbose=FALSE,nrounds=200)
  best_rounds <-append(best_rounds, cv.results_$best.nrounds)
}  
best_rounds
max(x)
```

```{r}
library(tidymodels)
install.packages("janitor")
stickelback_quang <- read_csv("https://www.dropbox.com/s/hpsn643fd8urph2/stickleback_fossil_vs_extant.csv?dl=1") %>% 
  janitor::clean_names() %>% 
  separate(univ_id, 
           into = c("lake", "body_type", "gender", "id"), 
           sep = "\\.",
           remove = FALSE)

stickelback_quang

extant_quan <- stickelback_quang %>%
  filter(lake != "Fos") %>% 
  select(gender, stl, ips, mds, mdf, maf, mav, mcv, mpt, ips_prop, 
         contains("_sc"), str_remove(names(stickelback_quang)[24:32], "_sc")) %>% 
  mutate(across(where(is_character), as_factor))
extant_quan

```
```{r}
my_data <- imputed_data_ng[1]
imputed_data_ng

my_data_new <- cbind(my_data, stick_clean[c("gender")])

my_data_extant <- my_data_new %>% 
  filter(gender == "Female" | gender=="Male") %>% 
  mutate(gender=as.factor(case_when(gender=="Female" ~ "F",
                   gender=="Male" ~ "M"

          )
         )
  )

```

```{r}
all_cores <- parallel::detectCores(logical = FALSE)
library(doParallel)

cl <- makePSOCKcluster(all_cores)
registerDoParallel(cl)
```

```{r}
set.seed(100)

split <- initial_split(my_data_extant, prop = 0.7, strata = gender)

train <- training(split)
test <- testing(split)

train_fold <- vfold_cv(train, v = 10, strata = gender)
train_fold$splits

xg_spec <- boost_tree(trees = tune(),
                      tree_depth = tune(), 
                      min_n = tune(), 
                      loss_reduction = tune(),
                      sample_size = tune(), 
                      mtry = tune(),
                      learn_rate = tune()) %>% 
  set_engine("xgboost") %>% 
  set_mode("classification")

xg_rec <- recipe(gender ~., data = train) %>% 
  step_dummy(all_nominal_predictors())

xg_rec$var_info

xg_wf <- workflow() %>% 
  add_model(xg_spec) %>% 
  add_recipe(xg_rec)

set.seed(105)
xg_grid <- grid_latin_hypercube(trees(),
                                tree_depth(),
                                min_n(),
                                loss_reduction(),
                                sample_size = sample_prop(),
                                finalize(mtry(), train),
                                learn_rate(),
                                size = 100)

xg_tune <- tune_grid(xg_wf,
                     resamples = train_fold,
                     grid = xg_grid,
                     control = control_grid(save_pred = TRUE))

xg_tune$.predictions
best_acc <- xg_tune %>% 
  select_best("accuracy")

best_acc

xg_final <- finalize_workflow(
  xg_wf,
  best_acc
)

xg_final

xg_last <- last_fit(xg_final, split)

xg_last %>% 
  collect_predictions() %>% 
  roc_curve(gender, .pred_F) %>% 
  ggplot(aes(1 - specificity, sensitivity)) +
  geom_abline() +
  geom_path() +
  coord_fixed()

xg_last %>% 
  collect_predictions() %>% 
  roc_auc(gender, .pred_F)

xg_last %>% 
  collect_predictions() %>% 
  accuracy(gender, .pred_class)
table(xg_last$.predictions[[1]]$.pred_class==xg_last$.predictions[[1]]$gender)

```

```{r}
library(tidymodels)
getImputedExtantData <- function(data,data2,num){
  data_out <- data[num]
  data_out <- cbind(data_out,data2[c("gender")])
  data_out <- data_out %>% 
    filter(gender == "Female" | gender=="Male") %>% 
    mutate(gender=as.factor(case_when(gender=="Female" ~ "F",
                   gender=="Male" ~ "M"

          )
         )
  )
  return(data_out)
}

getAccuracy  <- function(data){

  split <- initial_split(data, prop = 0.7, strata = gender)
  train <- training(split)
  test <- testing(split)

  train_fold <- vfold_cv(train, v = 10, strata = gender)

  xg_spec <- boost_tree(trees = tune(),
                        tree_depth = tune(), 
                        min_n = tune(), 
                        loss_reduction = tune(),
                        sample_size = tune(), 
                        mtry = tune(),
                        learn_rate = tune()) %>% 
    set_engine("xgboost") %>% 
    set_mode("classification")

  xg_rec <- recipe(gender ~., data = train) %>% 
    step_dummy(all_nominal_predictors())

  xg_wf <- workflow() %>% 
    add_model(xg_spec) %>% 
    add_recipe(xg_rec)
  xg_grid <- grid_latin_hypercube(trees(),
                                  tree_depth(),
                                  min_n(),
                                  loss_reduction(),
                                  sample_size = sample_prop(),
                                  finalize(mtry(), train),
                                  learn_rate(),
                                  size = 100)
  xg_tune <- tune_grid(xg_wf,
                       resamples = train_fold,
                       grid = xg_grid,
                       control = control_grid(save_pred = TRUE))

  best_acc <- xg_tune %>% 
    select_best("accuracy")

  xg_final <- finalize_workflow(
    xg_wf,
    best_acc
  )
  xg_last <- last_fit(xg_final, split)

  data_output <- xg_last %>% 
    collect_predictions() %>% 
    accuracy(gender, .pred_class)

  return(data_output)

}

imputed_data_tuned <- mixgb(data = non_gender, m=5, nrounds = 140)

new_list <- list()

for(i in 1:5){
  new_list[[i]] <- getAccuracy(getImputedExtantData(imputed_data_tuned,stick_clean,i))
}

new_list[[1]]
c <- c()

est/5
imputed_data_tuned

```