### 2023-01-23 stickle_imputation_error_bars.Rmd
pushing error bars plots for MICE imputation

---
title: "imputation_stickle"
author: "Akhil Ghosh"
output: html_document
---

```{r setup, include=FALSE}
library(mixgb)
library(mice)
library(tidyverse)
library(janitor)
library(missMethods)
library(ROCR)

```

DATA DICTIONARY
stl: standard length. A measure of size from tip of snout to end of tail. (mm)

ips: index pelvic score. a categorical measure of pelvic complexity based on the number of pelvic elements observed. 0-3 in fossils. 0-8 in extant fish.
ips.prop: ips/3 (fossils) or ips/8 (extant) to put ips on same scale from 0 to 1.
lps: length of a pelvic spine (if ips <3 (fossils) or ips < 8 (extant), then pelvic spines were considered evolutionarily lost, and were assigned lengths of 0)
tpg: pelvic girdle length (mm). (when missing (ips = 0), considered evolutionarily lost and assigned lengths of 0.)

mav: abdominal vertebrae number (count)
mcv: caudal vertebrae number (count)
mdf: dorsal fin ray number. number of bones in the dorsal fin (count)
maf: anal fin ray number (count)

mds: dorsal spine number (count)
ds1: length of the first dorsal spine (mm) (when missing, the first dorsal spine was considered evolutionarily lost, and assigned lengths of 0. when spines were broken assigned NA)
ds2: length of the second dorsal spine (mm) (when missing, the second dorsal spine was considered evolutionarily lost, and assigned lengths of 0. when spines were broken assigned NA)
ds3: length of the third dorsal spine (mm) (when missing, the third dorsal spine was considered evolutionarily lost, and assigned lengths of 0. when spines were broken assigned NA)

mpt: pterygiophore number (count). Number of armor bones on the "back". 
lpt: length of the pterygiophore anterior to the the third spine (mm)

ect: ectocoracoid length (mm)
cle: cleithrum length (mm)
pmx: premaxila length (mm)

```{r}
stickle <- read_csv('stickle_revamped.csv')

stick <- stickle %>% 
  mutate(gender = as.factor(case_when(
    grepl(".M.",univID,fixed=TRUE) ~ "Male",
    grepl(".F.",univID,fixed=TRUE) ~ "Female",
    grepl("Fos",univID,fixed=TRUE) ~ NA_character_
    )), ips = as.factor(ips)
  ) %>% 
  select(gender, everything())

stick

extant_stickle <- stick %>% filter(gender != "Fossil") %>% select(!univID)
extant_stickle
summary(extant_stickle)
```

```{r}

#Checking imputation accuracy/viability

M = 5
k = 5
stick_matrix <- matrix(NA, nrow=nrow(extant_stickle), ncol = M)
set.seed(12)
fold <- sample(1:k, size = nrow(extant_stickle), replace=TRUE)
for(i in 1:k){
extant_stickle_temp <- extant_stickle
extant_stickle_temp$gender[fold==i] <- NA
imputed_stick <- mice(extant_stickle_temp,m=M)
for(n in 1:M){

  complete_stick <- complete(imputed_stick, n)

  complete_stick$gender[fold==i]
  stick_matrix[fold==i, n] <- as.character(complete_stick$gender[fold==i])
}

}
pct_male <- apply(stick_matrix,1,function(x){
  mean(x == "Male")
})

imptest <- data_frame(truth = as.factor(extant_stickle$gender), pct_male)
roc_pred <- prediction(imptest$pct_male, imptest$truth)

auc_rocr <- performance(roc_pred, measure="auc")
plot(performance(roc_pred, measure="tpr", x.measure="fpr"))

auc_rocr@y.values[[1]]

```

```{r}
#Using MICE imputation to get imputed datasets
M = 5
stick_whole <- stick %>% select(-c(univID, ips))
summary(stick_whole)
imputed_stick_whole <- mice(stick_whole,m=5)
complete_stick_list <- list()
for(m in 1:M){
  complete_stick_list[[m]] <- complete(imputed_stick_whole, m)
}
```

```{r}
#getting time vector, adding it to imputed data
univ <- stick %>% 
  mutate(time = as.integer(sub(".*([0-9]{2})\\..*", "\\1",univID)))
times <- univ %>% drop_na(time) %>% pull(time)

times_vector <- c(rep(1,43)) %>% 
  append(21-times)
```

```{r}
#Initial grouping and summarizing
# temp_comp <- complete_stick_list[[1]][-c(1:367),]
# temp_comp$time <- times_vector
# 
# means_stl_time <- temp_comp %>% group_by(gender, time) %>% summarize(xbar = mean(stl), se = sqrt(var(stl)/n()), n())
# 
# 
# help(mice)
# temp_comp %>% select(gender, time, stl) %>% filter(time==17)
# 
# 
# with(temp_)
# 
# means_stl_time %>% ggplot(aes(x = time, y = xbar, col = gender))+ geom_point() +geom_line()
# 
# table(complete_stick_list[[1]][c(1:367),]$gender)
# 
# table(complete_stick_list[[1]][-c(1:367),]$gender)

```

```{r}
#doing the grouping and summarizing for all imputed datasets, looking specifically at mean standard length for each group

temp_comp_lists <- list()
est_list <- list()
M = 5
for(m in 1:M){

temp_comp_lists[[m]] <- complete_stick_list[[m]][-c(1:367),]
temp_comp_lists[[m]]$time <- times_vector

est_list[[m]] <- temp_comp_lists[[m]] %>% group_by(gender, time) %>% summarize(xbar = mean(stl), se = sqrt(var(stl)/n()), n())
est_list[[m]]$imp = m
}

# stacking the datasets
combined_est <- do.call(rbind,est_list)
combined_est
```

```{r}
#Calculating error bars of combined estimates

combined_est_errors <- combined_est %>% group_by(gender, time) %>% summarize(x_bar_bar  = mean(xbar), B_m = var(xbar), W_m = mean(se^2), se_xbar_bar = sqrt(W_m+(1+1/5)*(B_m)), r = (1+1/5)*B_m/W_m, U = (5-1)*(1+(1/r)^2), t_val = qt(0.05/2, df = U, lower.tail = FALSE), uci = x_bar_bar + t_val * se_xbar_bar, lci = x_bar_bar - t_val * se_xbar_bar, t_se = t_val * se_xbar_bar)

combined_est_errors %>% select(gender, time,x_bar_bar, lci, uci)

```

```{r}
#ggplot of the data for

ggplot(data= combined_est_errors, aes(x = time, y = x_bar_bar, col = gender))+geom_point()+geom_line() + geom_errorbar(aes(ymin = lci, ymax = uci, group = gender),
     width = 0.2)+ labs(x = "Time", y = "Mean of imputed mean standard length values (STL)")

```

```{r}
#Doing it for cle variable this time
temp_comp_lists_cle <- list()
est_list_cle <- list()
M = 5
for(m in 1:M){

temp_comp_lists_cle[[m]] <- complete_stick_list[[m]][-c(1:367),]
temp_comp_lists_cle[[m]]$time <- times_vector

est_list_cle[[m]] <- temp_comp_lists_cle[[m]] %>% group_by(gender, time) %>% summarize(xbar = mean(cle), se = sqrt(var(cle)/n()), n())
est_list_cle[[m]]$imp = m
}

est_list_cle

combined_est_cle <- do.call(rbind,est_list_cle)

#Calculating error bars of combined estimates

combined_est_errors_cle <- combined_est_cle %>% group_by(gender, time) %>% summarize(x_bar_bar  = mean(xbar), B_m = var(xbar), W_m = mean(se^2), se_xbar_bar = sqrt(W_m+(1+1/5)*(B_m)), r = (1+1/5)*B_m/W_m, U = (5-1)*(1+(1/r)^2), t_val = qt(0.05/2, df = U, lower.tail = FALSE), uci = x_bar_bar + t_val * se_xbar_bar, lci = x_bar_bar - t_val * se_xbar_bar, t_se = t_val * se_xbar_bar)

combined_est_errors_cle %>% select(gender, time,x_bar_bar, lci, uci)

ggplot(data= combined_est_errors_cle, aes(x = time, y = x_bar_bar, col = gender))+geom_point()+geom_line() + geom_errorbar(aes(ymin = lci, ymax = uci),
     width = 0.2) + labs(x = "Time", y = "Mean of imputed mean CLE")

```