### 2023-04-04 mixgb_errors.Rmd

---
title: "mixgb_imputation"
author: "Akhil Ghosh"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE}
library(mice)
library(tidyverse)
library(janitor)
library(missMethods)
library(ROCR)

```

```{r}
stickle <- read_csv('stickle_revamped.csv')

stick <- stickle %>% 
  mutate(gender = as.factor(case_when(
    grepl(".M.",univID,fixed=TRUE) ~ "Male",
    grepl(".F.",univID,fixed=TRUE) ~ "Female",
    grepl("Fos",univID,fixed=TRUE) ~ NA_character_
    )), ips = as.factor(ips)
  ) %>% 
  select(gender, everything())

stick_whole <- stick %>% select(-c(univID, ips))

extant_stickle <- stick %>% filter(gender != "Fossil") %>% select(!univID)
extant_stickle <- extant_stickle %>% select(!ips)

summary(extant_stickle)
```

```{r}
k = 5
M = 5

#generate 30% MCAR for gender on extant to find best hyperparams to train mixgb
#this is most likely data leakage of some form, but I might as well try things and then have Greg tell me I'm an idiot later
extant_stickle_MAR <- delete_MCAR(extant_stickle, 1, "gender")

mixgb_cv(extant_stickle, response = "gender")
#seems like best rounds is rarely, if ever, above 15-20.
#tracks with my previous forays into doing this

#basically, the issue I'm having is how should I "train" the mixgb model without data leakage.
#maybe since its used for imputation, and not prediction it really doesn't matter
#need to consult greg
```

```{r}
library(xgboost)
library(caret)
library(caTools)
library(tidymodels)

#how about we do a grid search to train a xgboost model to predict gender of extants, and use those hyperparameters in the mixgb model for imputation?
#worth a try

extant_stickle

set.seed(2023) # For reproducibility
split <- initial_split(extant_stickle, prop = 0.7)
train <- training(split)
test <- testing(split)
```

```{r}
#recipe to pre-process data

set.seed(2023)
split <- initial_split(extant_stickle, prop = 0.7, strata = gender)
train <- training(split)
test <- testing(split)
library(x)
dtrain <- xgb.DMatrix(as.matrix(train[,-1]), label = (as.numeric(train$gender)-1))
dtest <- xgb.DMatrix(as.matrix(test[, -1]), label = (as.numeric(test$gender)-1))
```

```{r}

params <- list(
  booster = "gbtree",
  objective = "binary:logistic",
  eta = 0.3,
  max_depth = 6,
  min_child_weight = 1,
  subsample = 1,
  colsample_bytree = 1,
  eval_metric = "error",
  n_thread=3
)

# Train the xgboost model
xgb_model <- xgb.train(
  params = params,
  data = dtrain,
  nrounds = 100,
  watchlist = list(train = dtrain, test = dtest),
  print_every_n = 10
)

# Make predictions on the test set
pred <- predict(xgb_model, dtest)

# Calculate the accuracy
table(test$gender)
accuracy <- mean((pred > 0.7) == (as.numeric(test$gender)-1))
cat("Accuracy:", accuracy, "\n")
```

```{r}

folds <- vfold_cv(train, v = 5, strata=gender)
```

```{r}
xg_spec <- boost_tree(trees = tune(),
                      tree_depth = tune(), 
                      min_n = tune(), 
                      loss_reduction = tune(),
                      sample_size = tune(), 
                      mtry = tune(),
                      learn_rate = tune()) %>% 
  set_engine("xgboost") %>% 
  set_mode("classification")

xg_rec <- recipe(gender ~., data = train) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_impute_median(all_numeric_predictors())

xg_wf <- workflow() %>% 
  add_model(xg_spec) %>% 
  add_recipe(xg_rec)

# tune
set.seed(105)
xg_grid <- grid_latin_hypercube(trees(),
                                tree_depth(),
                                min_n(),
                                loss_reduction(),
                                sample_size = sample_prop(),
                                finalize(mtry(), train),
                                learn_rate(),
                                size = 50)

xg_tune <- tune_grid(xg_wf,
                     resamples = folds,
                     grid = xg_grid,
                     control = control_grid(save_pred = TRUE))
best_acc <- xg_tune %>% 
  select_best("accuracy")
```

```{r}
library(future)

rec <- recipe(gender ~ ., data = train) %>%
  step_dummy(all_nominal_predictors()) %>% 
  step_impute_median(all_numeric()) %>% 
  step_scale(all_numeric())

xg_spec <- boost_tree(trees = tune(),
                      tree_depth = tune(), 
                      min_n = tune(), 
                      loss_reduction = tune(),
                      sample_size = tune(), 
                      mtry = tune(),
                      learn_rate = tune()) %>%   set_engine("xgboost") %>% set_mode("classification")

xg_wf <- workflow() %>% 
  add_model(xg_spec) %>% 
  add_recipe(rec)

#grid search
xg_grid <- grid_latin_hypercube(trees(),
                                tree_depth(),
                                min_n(),
                                max_n(),
                                loss_reduction(),
                                sample_size = sample_prop(),
                                finalize(mtry(), train),
                                learn_rate(),
                                size = 50)

best_model <- select_best(xg_tune, metric = "accuracy")

```

```{r}
library(doParallel)
cl <- makeCluster(8)
registerDoParallel(cl)

xg_tune <- tune_grid(xg_wf,
                     resamples = folds,
                     grid = xg_grid,
                     control = control_grid(save_pred = TRUE, parallel_over = "everything", allow_par = T))

best_acc <- select_best(xg_tune, metric = "accuracy")
```

```{r}
xg_final <- finalize_workflow(
  xg_wf,
  best_acc
)

xg_final

xg_last <- last_fit(xg_final, split)

xg_last %>% 
  collect_predictions() %>% 
  accuracy(gender, .pred_class)

xg_last %>% 
  collect_predictions() %>% 
  roc_curve(gender, .pred_Female) %>% 
  ggplot(aes(1 - specificity, sensitivity)) +
  geom_abline() +
  geom_path() +
  coord_fixed()

xg_last %>% 
  collect_predictions() %>% 
  roc_auc(gender, .pred_Male)

xg_last %>% 
  collect_predictions() %>% 
  accuracy(gender, .pred_class)

```

```{r}
stickle_wf <- workflow() %>%
  add_recipe(recipe) %>%
  add_model(xgbtree())

```

```{r}
library(mixgb)

```

```{r}
#okay, onto the real thing here
#Looking to validate how imputation is doing against reality
M=5
k = 5
stick_matrix <- matrix(NA, nrow=nrow(extant_stickle), ncol = M)

set.seed(2023) # seed used for 86% was 12

fold <- sample(1:k, size = nrow(extant_stickle), replace=TRUE)
for(i in 1:k){
extant_stickle_temp <- extant_stickle
extant_stickle_temp$gender[fold==i] <- NA

#imputed_stick <- mixgb(extant_stickle_temp,nrounds = 100, m = 5, xgb.params = list(max_depth = 12, eta = 0.014311225124087, gamma = 0,min_child_weight = 1,subsample = 1, colsample_bytree = 1, colsample_bylevel = 1, colsample_bynode = 1,tree_method = "auto", gpu_id = 0, predictor = "auto"))

imputed_stick <- mixgb(extant_stickle_temp, nrounds = 100, m=5)
#all that tuning for it to be a worse imputer....
#whatever

for(n in 1:M){

  complete_stick <- imputed_stick[[n]]

  complete_stick$gender[fold==i]
  stick_matrix[fold==i, n] <- as.character(complete_stick$gender[fold==i])
}

}

pct_male <- apply(stick_matrix,1,function(x){
  mean(x == "Male")
})

imptest <- data_frame(truth = as.factor(extant_stickle$gender), pct_male)
roc_pred <- prediction(imptest$pct_male, imptest$truth)

auc_rocr <- performance(roc_pred, measure="auc")
plot(performance(roc_pred, measure="tpr", x.measure="fpr"))

auc_rocr@y.values[[1]]

```

```{r}
#Using MixGB imputation to get imputed datasets
M = 5
stick_whole <- stick %>% select(-c(univID, ips))
imputed_stick_whole <- mixgb(stick_whole,m=5, nrounds=100)
imputed_stick_whole[[1]]
```

```{r}
univ <- stick %>% 
  mutate(time = as.integer(sub(".*([0-9]{2})\\..*", "\\1",univID)))
times <- univ %>% drop_na(time) %>% pull(time)

times_vector <- c(rep(1,43)) %>% 
  append(21-times)
```

```{r}
#again issues with not having enough males :/
#what to do...

temp_comp_lists <- list()
est_list <- list()
M = 5
for(m in 1:M){

temp_comp_lists[[m]] <- imputed_stick_whole[[m]][-c(1:367),]
temp_comp_lists[[m]]$time <- times_vector

est_list[[m]] <- temp_comp_lists[[m]] %>% group_by(gender, time) %>% summarize(xbar = mean(stl), se = sqrt(var(stl)/n()), n())
est_list[[m]]$imp = m
}

# stacking the datasets
combined_est <- do.call(rbind,est_list)
combined_est
```

```{r}
#Calculating error bars of combined estimates

combined_est_errors <- combined_est %>% group_by(gender, time) %>% summarize(x_bar_bar  = mean(xbar), B_m = var(xbar), W_m = mean(se^2), se_xbar_bar = sqrt(W_m+(1+1/5)*(B_m)), r = (1+1/5)*B_m/W_m, U = (5-1)*(1+(1/r)^2), t_val = qt(0.05/2, df = U, lower.tail = FALSE), uci = x_bar_bar + t_val * se_xbar_bar, lci = x_bar_bar - t_val * se_xbar_bar, t_se = t_val * se_xbar_bar)

combined_est_errors %>% select(gender, time,x_bar_bar, lci, uci)

```

```{r}
ggplot(data= combined_est_errors, aes(x = time, y = x_bar_bar, col = gender))+geom_point()+geom_line() + geom_errorbar(aes(ymin = lci, ymax = uci, group = gender),
     width = 0.2)+ labs(x = "Time", y = "Mean of imputed mean standard length values (STL)")

```