### 2023-05-01 R/Stickle_plots.Rmd

---
title: "stickle_plots"
author: "Akhil Ghosh"
date: "2022-10-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(mice)
library(tidyverse)
library(tidymodels)

library(mitools)
library(miceadds)

library(janitor)
library(furrr)
library(doParallel)
library(mixgb)
```

```{r}
stickle <- read_csv('stickle_revamped.csv')

stick <- stickle %>% 
  mutate(gender = as.factor(case_when(
    grepl(".M.",univID,fixed=TRUE) ~ "Male",
    grepl(".F.",univID,fixed=TRUE) ~ "Female",
    grepl("Fos",univID,fixed=TRUE) ~ NA_character_
    )), ips = as.factor(ips)
  ) %>% 
  select(gender, everything())
```

```{r}
set.seed(205)

stick_clean <- stick %>% 
  select(!c(univID, ips))

stick_clean <- data_clean(stick_clean)
stick_imputed <- mixgb(data = stick_clean, m = 10, nrounds = 20)
stick_imputed
```

```{r}
mice(stick_clean, )
as <- mice(stick_clean, method = "pmm")
as <- mice(stick_clean, method = "rf")
as$imp

help(datalist2mids)
```

```{r}
#getting time variable
datalist2mids(stick_$imputations)
univ <- stick %>% 
  mutate(time = as.integer(sub(".*([0-9]{2})\\..*", "\\1",univID)))
times <- univ %>% drop_na(time) %>% pull(time)

times_vector <- c(rep(1,43)) %>% 
  append(21-times)

```

```{r}
add_time<- function(data){
  data %>% 
  slice(368:1181) %>%
  mutate(time = times_vector) %>% 
  select(time, everything())
}

means_by_time <- function(data){
  data %>% group_by(time, gender) %>% 
  summarize(across(where(is.numeric),mean))
}

mean_diff <- function(data){
  data %>% group_by(time, gender) %>% 
  summarize(across(where(is.numeric),mean)) %>% 
  summarize(across(where(is.numeric),diff))

}

mean_diff2 <- function(data){
  data %>% 
  summarize(across(where(is.numeric),diff))
}
```

```{r}
set.seed(205)
stick_ <- imputationList(stick_imputed)

final <- with(imputationList(with(imputationList(stick_imputed), fun=add_time)),fun=means_by_time)

mean_differences <- with(imputationList(final[-c(6:10)]),fun=mean_diff2)
mean_differences

fin <- (mean_differences[[1]]+mean_differences[[2]]+mean_differences[[3]]+mean_differences[[4]]+mean_differences[[5]])/5
fin

fin %>% ungroup() %>%
  select(where(is.numeric)&!time) %>% colnames() %>% 
  map( ~ {ggplot(fin, aes(x = time,y=.data[[.]])) + geom_line()})

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.