### 2023-05-01 R/rf_imputation.Rmd

---
title: "mice_rf"
author: "Akhil Ghosh"
output: html_document
---

```{r setup, include=FALSE}
library(mice)
library(tidyverse)
library(janitor)
library(missMethods)
library(ROCR)
```

```{r setup, include=FALSE}
stickle <- read_csv('stickle_revamped.csv')

stick <- stickle %>% 
  mutate(gender = as.factor(case_when(
    grepl(".M.",univID,fixed=TRUE) ~ "Male",
    grepl(".F.",univID,fixed=TRUE) ~ "Female",
    grepl("Fos",univID,fixed=TRUE) ~ NA_character_
    )), ips = as.factor(ips)
  ) %>% 
  select(gender, everything())

stick_whole <- stick %>% select(-c(univID, ips))

extant_stickle <- stick %>% filter(gender != "Fossil") %>% select(!univID)
extant_stickle <- extant_stickle %>% select(!ips)

mice(extant_stickle, method="rf")

```

```{r}
library(mice)

```

```{r}
library(missForest)
summary(extant_stickle)
test <- extant_stickle
missing_cols <- c(2:7,9:27)
extant_stickle_MAR <- delete_MCAR(extant_stickle, 0.3, "gender")
imputed_data <- missForest(extant_stickle_MAR)
test[, missing_cols] <- imputed_data$ximp[, missing_cols]
```

```{r}
extant_stickle_df <- as.data.frame(extant_stickle)
extant_stickle_impo <- missForest(extant_stickle_df)

new_data <- extant_stickle_impo$ximp
new_data

doParallel::registerDoParallel(cores = 8) # set based on number of CPU cores
doRNG::registerDoRNG(seed = 2023)
new_data <- missForest(extant_stickle_df, parallelize = 'forests')$ximp

```

```{r}
M = 5
k = 5
mice(extant_stickle, method = 'rf', rfpackage = "randomForest")

stick_matrix <- matrix(NA, nrow=nrow(extant_stickle), ncol = M)
set.seed(2023)
fold <- sample(1:k, size = nrow(extant_stickle), replace=TRUE)
for(i in 1:k){
extant_stickle_temp <- extant_stickle
extant_stickle_temp$gender[fold==i] <- NA
extant_stickle_temp_df <- as.data.frame(extant_stickle_temp)
doParallel::registerDoParallel(cores = 8) # set based on number of CPU cores
doRNG::registerDoRNG(seed = 2023)

imputed_stick <- missForest(extant_stickle_temp_df, parallelize = 'variables')$ximp

for(n in 1:M){

  imputed_stick$gender[fold==i]
  stick_matrix[fold==i, n] <- as.character(imputed_stick$gender[fold==i])
}

}

pct_male <- apply(stick_matrix,1,function(x){
  mean(x == "Male")
})

imptest <- data_frame(truth = as.factor(extant_stickle$gender), pct_male)
roc_pred <- prediction(imptest$pct_male, imptest$truth)

auc_rocr <- performance(roc_pred, measure="auc")
plot(performance(roc_pred, measure="tpr", x.measure="fpr"))

auc_rocr@y.values[[1]]
```

```{r}
mice(stick, method='rf')
```