### 2023-06-23 R/StickleImputationCode.Rmd
commiting R code for stickle imputation process and committing the updated imputations

---
title: "StickleImputationCode"
author: "Akhil Ghosh"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mice)
```

```{r}
setwd('/Users/akhilghosh/Desktop/RStudioProjects2022Fall/SticklebackProject/RawData_ReadOnly')
stickle <- read_csv("AO.sexualdimorphismPruned.sc.forAnalysis_forMICE_Modeling_20June2023.csv")

summary(stickle)
stick <- stickle %>% 
  mutate(gender = as.factor(case_when(
    grepl(".M.",univID,fixed=TRUE) ~ "Male",
    grepl(".F.",univID,fixed=TRUE) ~ "Female",
    grepl("Fos",univID,fixed=TRUE) ~ NA_character_
    ))
  ) %>% 
  select(gender, everything())

```

```{r}
set.seed(0314)
imputations = 100
stick_whole <- stick %>% select(-c(univID))
summary(stick_whole)
imputed_stick_whole <- mice(stick_whole,m=imputations)
complete_stick_list <- list()
for(m in 1:imputations){
  complete_stick_list[[m]] <- complete(imputed_stick_whole, m)
}
```

```{r}
univ <- stick %>% 
  mutate(time = as.integer(sub(".*([0-9]{2})\\..*", "\\1",univID)))
times <- univ %>% drop_na(time) %>% pull(time)

times_vector <- c(rep(1,43)) %>% 
  append(21-times)
```

```{r}
csv_list <-list()
for(i in 1:imputations){
csv_list[[i]] <- complete_stick_list[[i]][-c(1:367),] %>% mutate(imp = rep(i, 814), time = times_vector)
}

combined_csv<- do.call(rbind,csv_list) %>% select(imp, time,gender, everything())
```