### 2024-09-30 manuscript/paper.tex
Update manuscript with posterior and Kendall's Tau

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{setspace} \setstretch{1.15} \usepackage{float} \floatplacement{figure}{t}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  colorlinks=true,
  linkcolor={cyan},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={cyan},
  pdfcreator={LaTeX via pandoc}}

\title{~\Large Increased sexual dimorphism evolves in a fossil
stickleback following ecological release from fish piscivores}
\author{\large Matthew Stuart\(^{1,2}\), Allison Ozark\(^{3}\), Raheyma
Siddiqui\(^3\), Akhil Ghosh\(^{1,2}\)\\
\large Samantha Swank\(^3\), Michael A. Bell\(^4\), Gregory J.
Matthews\(^{1,2}\), and Yoel E. Stuart\(^{3,+}\)\\
\vspace{-1.1mm}\\
\large \(^1\) Department of Mathematics and Statistics, Loyola
University Chicago, Chicago, IL, USA \vspace{-1.1mm}\\
\large \(^2\) Center for Data Science and Consulting, Loyola University
Chicago, Chicago, IL, USA \vspace{-1.1mm}\\
\large \(^3\) Department of Biology, Loyola University Chicago, Chicago,
IL, USA \vspace{-1.1mm}\\
\large \(^4\) University of California Museum of Paleontology, Berkeley,
CA, USA \vspace{-1.1mm}\\
\large \(^+\) Corresponding:
\href{mailto:ystuart@luc.edu}{\nolinkurl{ystuart@luc.edu}}
\vspace{-1.1mm}}
\date{}

\begin{document}
\maketitle
\begin{abstract}
Everyone loves the stickle \vspace{2mm}\\
\emph{Keywords}: Stickle
\end{abstract}

\newcommand{\iid}{\overset{iid}{\sim}}

\newpage

\hypertarget{sec:intro}{%
\section{Introduction}\label{sec:intro}}

Ecological release theory suggests that a population's niche should
change when important species interactions like resource competition or
predation are relaxed or removed (reviewed in Herrmann, Stroud, and
Losos (\protect\hyperlink{ref-Herrmannetal2021}{2021})). Removal is
posited to create ecological opportunity---i.e., aspects of the niche
become newly accessible, and the focal population shifts and/or expands
its resource use new resources (Parent and Crespi
(\protect\hyperlink{ref-ParentandCrespi2009}{2009}); Herrmann, Stroud,
and Losos (\protect\hyperlink{ref-Herrmannetal2021}{2021})). Ecological
release may then be followed by adaptive morphological evolution as
traits change to reflect the new niche (Parent and Crespi
(\protect\hyperlink{ref-ParentandCrespi2009}{2009}); Herrmann, Stroud,
and Losos (\protect\hyperlink{ref-Herrmannetal2021}{2021})).

For example, a population undergoing ecological release can experience
disruptive selection on males and females stemming from intraspecific
competition over newly accessible resources (Bolnick and Doebeli
(\protect\hyperlink{ref-BolnickandDoebeli2003}{2003}); Bolnick and Lau
(\protect\hyperlink{ref-BolnickandLau2008}{2008}); Cooper, Gilman, and
Boughman (\protect\hyperlink{ref-Cooperetal2011}{2011})). This is
predicted to result in intersexual divergence in habitat use and
associated phenotypes (Schoener
(\protect\hyperlink{ref-Schoener1968}{1968}); Shine
(\protect\hyperlink{ref-Shine1989}{1989}); Bolnick and Doebeli
(\protect\hyperlink{ref-BolnickandDoebeli2003}{2003}); Butler and Losos
(\protect\hyperlink{ref-Butleretal2007}{2007}); Bolnick and Lau
(\protect\hyperlink{ref-BolnickandLau2008}{2008}); Cooper, Gilman, and
Boughman (\protect\hyperlink{ref-Cooperetal2011}{2011}); but see Stuart
et al. (\protect\hyperlink{ref-Stuartetal2021}{2021}); Blain
(\protect\hyperlink{ref-Blain2022}{2022})). Such competition-driven
``intraspecific character displacement'' between the sexes is therefore
one explanation for the evolution of sexual dimorphism (Pfennig and
Pfennig (\protect\hyperlink{ref-PfennigandPfennig2012}{2012}); De Lisle
and Rowe (\protect\hyperlink{ref-DeLisleandRowe2015}{2015}); De Lisle
and Rowe (\protect\hyperlink{ref-DeLisleandRowe2017}{2017}), De Lisle,
Paiva, and Rowe (\protect\hyperlink{ref-DeLisleetal2018}{2018})).

Much of the theory and empirical data for character displacement between
the sexes is based on release from resource competition specifically
(e.g., Bolnick and Doebeli
(\protect\hyperlink{ref-BolnickandDoebeli2003}{2003}); Cooper, Gilman,
and Boughman (\protect\hyperlink{ref-Cooperetal2011}{2011}); Pfennig and
Pfennig (\protect\hyperlink{ref-PfennigandPfennig2012}{2012}); De Lisle
and Rowe (\protect\hyperlink{ref-DeLisleandRowe2015}{2015}); De Lisle,
Paiva, and Rowe (\protect\hyperlink{ref-DeLisleetal2018}{2018})).
However, release from predation might also drive the evolution of
increased sexual dimorphism because an absence of predators should
generate ecological opportunity (Reimchen and Nosil
(\protect\hyperlink{ref-ReimchenandNosil2004}{2004}); Parent and Crespi
(\protect\hyperlink{ref-ParentandCrespi2009}{2009}); Herrmann, Stroud,
and Losos (\protect\hyperlink{ref-Herrmannetal2021}{2021})).

Here, we tested the prediction that a release from predation results in
the evolution of increased sexual dimorphism using a well-preserved,
finely-resolved sequence of a fossil threespine stickleback fish
(Gasterosteus doryssus). The sequence in this depositional environment
is comprised of two lineages (Bell
(\protect\hyperlink{ref-Bell2009}{2009}); Cerasoni, Bell, and Stuart
(\protect\hyperlink{ref-Cerasonietal2024}{2024})). Lineage I was a
low-armor form with zero to one dorsal spines and highly reduced
pelvises, on average. Lineage I lasted for at least 93,000 years before
it was replaced by a second lineage, suddenly, likely on the order of
years (Bell, Baumgartner, and Oslon
(\protect\hyperlink{ref-Belletal1985}{1985}); Stuart et al.~unpublished
data). Lineage II appeared in the depositional environment fully
armored, with complete pelvic girdles, two pelvic spines, and three
dorsal spines (Bell, Travis, and Blouw
(\protect\hyperlink{ref-Belletal2006}{2006}); Stuart, Travis, and Bell
(\protect\hyperlink{ref-Stuartetal2020}{2020})). It then immediately
began evolving adaptive reduction in its armor traits (Hunt, Bell, and
Travis (\protect\hyperlink{ref-Huntetal2008}{2008}); Stuart, Travis, and
Bell (\protect\hyperlink{ref-Stuartetal2020}{2020})) until it reached
the same low-armor state previously held by lineage I.

The observation of 93,000 years of low-armor stasis in lineage I (Bell,
Baumgartner, and Oslon (\protect\hyperlink{ref-Belletal1985}{1985})) and
the observation of rapid evolution of armor loss by lineage II, both
suggest that this depositional environment lacked piscivorous fish like
trout and other salmonids known to prey on modern threespine
stickleback. Armor presence in extant threespine stickleback
(Gasterosteus aculeatus) correlates strongly with the presence of
vertebrate piscivores; populations with less predation pressure
typically have less armor (Reimchen
(\protect\hyperlink{ref-Reimchen1994}{1994}), Reimchen and Nosil
(\protect\hyperlink{ref-ReimchenandNosil2004}{2004}); Bell et al.
(\protect\hyperlink{ref-Belletal1993}{1993}); Roesti et al.
(\protect\hyperlink{ref-Roestietal2023}{2023})). In our paleolake basin,
only three fossil trout have been found in the same section of rock that
has revealed \textgreater20,000 threespine stickleback fossils as well
as occasional killifish (Fundulus nevadensis) (Bell
(\protect\hyperlink{ref-Bell2009}{2009}); Cerasoni, Bell, and Stuart
(\protect\hyperlink{ref-Cerasonietal2024}{2024})). Thus, we reconstruct
an evolutionar history in which it is likely that lineage II migrated
from a nearby paleolake basin that had predators since it was armored
when it arrived (Bell (\protect\hyperlink{ref-Bell2009}{2009});
Cerasoni, Bell, and Stuart
(\protect\hyperlink{ref-Cerasonietal2024}{2024})). Lineage II then
experienced release from predators in the focal paleolake basin,
generating the initial conditions of evolutionary models predicting the
evolution of increased sexual dimorphism following ecological release.
We test this prediction here.

A major challenge in the study of fossilized sexual dimorphism is
assigning sex to individual specimens in the first place (Hone and
Mallon (\protect\hyperlink{ref-HoneandMallon2017}{2017}); Mallon
(\protect\hyperlink{ref-Mallon2017}{2017}); Saitta et al.
(\protect\hyperlink{ref-Saittaetal2020}{2020})). This typically cannot
be done directly, except for taxa whose sexes are distinguished by the
presence or absence of sex-specific characters that preserve well.
Instead, paleobiologists often resort to statistical detection of sex
and sexual dimorphism, including tests for normality and bimodality in
trait distributions (e.g., Mallon
(\protect\hyperlink{ref-Mallon2017}{2017})), mixture modeling (e.g.,
Mallon (\protect\hyperlink{ref-Mallon2017}{2017})), divergence in growth
curves (e.g., Saitta et al.
(\protect\hyperlink{ref-Saittaetal2020}{2020})), and emphasis on effect
size statistics rather than significance testing (e.g., Saitta et al.
(\protect\hyperlink{ref-Saittaetal2020}{2020})). However, dimorphic
signal can be masked by noise introduced by factors both biological and
artifactual, including extended growth during ontogeny, age-based bias
in survivorship, small sample sizes, time averaging, and taphonomic bias
(Godfrey, Lyon, and Sutherland
(\protect\hyperlink{ref-Godfreyetal1993}{1993}); Koscinski and
Pietraszewski
(\protect\hyperlink{ref-KoscinskiandPietraszewski2004}{2004}); Hone and
Mallon (\protect\hyperlink{ref-HoneandMallon2017}{2017}); reviewed in
Mallon (\protect\hyperlink{ref-Mallon2017}{2017}); reviewed in Saitta et
al. (\protect\hyperlink{ref-Saittaetal2020}{2020})). Thus, the best
approach to sex classification is likely one of total evidence (Saitta
et al. (\protect\hyperlink{ref-Saittaetal2020}{2020})), including
comparison to closely-related, extant species of known sex (e.g., Hone
and Mallon (\protect\hyperlink{ref-HoneandMallon2017}{2017}); Saitta et
al. (\protect\hyperlink{ref-Saittaetal2020}{2020})).

For our study, we inferred sex in G. doryssus fossils by comparing
multivariate morphological trait data from lineage II samples to the
same multivariate trait set collected from multiple populations in the
closely-related, extant Threespine Stickleback species complex
(Gasterosteus aculeatus). Crucially, we determined G. aculeatus sex
directly via dissection and/or PCR genotyping. This enabled us to use
Multiple Imputation by Chained Equations (MICE; Buuren and
Groothuis-Oudshoorn (\protect\hyperlink{ref-MICE}{2011})) to build a
predictive multiple imputation algorithm (Little and Rubin
(\protect\hyperlink{ref-little2002statistical}{2002})) based on the
multivariate morphology of G. aculeatus individuals of known sex. We
applied this algorithm to the fossil data to impute individual sex based
on morphology, treating fossil sex as a missing variable. With sex
assigned to fossil specimens, we fit a modified Ornstein-Uhlenbeck (OU)
(Uhlenbeck and Ornstein (\protect\hyperlink{ref-OUProcess}{1930})) model
using a Bayesian framework that accounts for uncertainty in sex
classification to test for evolution of sexual dimorphism in each trait
over \textasciitilde16,000 years of lineage II.

\hypertarget{sec:data}{%
\section{Data}\label{sec:data}}

\hypertarget{fossil-specimen-data}{%
\subsection{Fossil Specimen Data}\label{fossil-specimen-data}}

We used Gasterosteus doryssus data that were previously reported by
Stuart, Travis, and Bell (\protect\hyperlink{ref-Stuartetal2020}{2020}),
Voje, Bell, and Stuart (\protect\hyperlink{ref-Vojeetal2022}{2022}), and
Siddiqui et al. (\protect\hyperlink{ref-Siddiquietal2024}{2024}).
Briefly, the data were collected from fossil Series K from Quarry D
(Cerasoni, Bell, and Stuart
(\protect\hyperlink{ref-Cerasonietal2024}{2024})), dug from an open pit
diatomite mine at 9.526° N, 119.094° W, near Hazen, Nevada, USA. Series
K consisted of 18 samples taken at \textasciitilde1000-year intervals,
and mean sample times span \textasciitilde16,363 years. Fish from series
K were measured for 16 ecomorphological traits related to armor,
swimming, and feeding (Table 1). Series K started at the previously
documented horizon when lineage I was replaced by lineage II. The tempo
and mode of lineage II armor reduction during this sequence suggests
adaptive evolution by natural selection (Hunt, Bell, and Travis
(\protect\hyperlink{ref-Huntetal2008}{2008})), and we focus on the
multivariate evolution of sexual dimorphism by this second lineage.

The lineage II fossil data consist of 814 specimens of unknown sex
sampled across the 18 K series samples. Figure 1 shows the sample size
for each of the 18 samples. There are at least 22 specimens in each
sample with a high of 67 specimens in sample 7.

THIS FIGURE DOESN'T SHOW MUCH. LET'S CUT FOR SPACE.

\begin{figure}

{\centering \includegraphics{paper_files/figure-latex/unnamed-chunk-2-1}

}

\caption{Fossil sample size for each time period}\label{fig:unnamed-chunk-2}
\end{figure}

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.3492}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1905}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4603}}@{}}
\caption{Traits and trait descriptions. `sc' denotes size correction of
trait against standard length. Names of bones follow Bowne
(\protect\hyperlink{ref-Bowne1994}{1994}) unless otherwise
noted.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Trait Name
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Trait Code
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Trait Description
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Trait Name
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Trait Code
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Trait Description
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Standard Length & \texttt{stl} & Distance from anterior tip of
premaxilla to posterior end of last vertebra (hypural plate) \\
Dorsal Spine & \texttt{mds} & Number of dorsal spines from 0 to 3 \\
Dorsal Fin Ray & \texttt{mdf} & Number of bones in the dorsal fin
posterior to the third dorsal spine (i.e., soft dorsal fin rays) \\
Anal Fin Ray & \texttt{maf} & Number of bones in the anal fin posterior
to the anal spine (i.e., soft dorsal fin rays) \\
Abdominal Vertebra & mav & Number of vertebrae anterior to the first
vertebra contacting an anal fin pterygiophore (Aguirre et al.~2014) \\
Caudal Vertebra & \texttt{mcv} & Number of vertebrae posterior to and
including the first vertebra contacting an anal fin pterygiophore
(Aguirre et al.~2014) \\
Pterygiophore number & \texttt{mpt} & Number of pterygiophores anterior
to but excluding the pterygiophore under the third dorsal spine, which
is immediately anterior to and contiguous with the dorsal fin \\
Pelvic Spine length & \texttt{lps.sc} & Length from the base of one
pelvic spine above its articulation with the pelvic girdle to its distal
tip \\
Ectocoracoid & \texttt{ect.sc} & Length between the anterior and
posterior tips of the shoulder girdle base (i.e., ectocoracoid) \\
Pelvic Girdle & \texttt{tpg.sc} & Length between the anterior to
posterior tips along midline. If vestigial, the sum of longest
anterior-posterior axis for the vestiges \\
Cleithrum length & \texttt{cle.sc} & Length from free dorsal tip to
ventral tip of the cleithrum on the anterior margin of the shoulder
girdle (i.e., cleithrum) \\
Premaxilla & \texttt{pmx.sc} & Length from the anterior tip of the
premaxilla to the distal tip of the ascending process of the
premaxilla \\
Dorsal Spine & \texttt{Ds\#.sc\#\ =\ 1,2,or\ 3} & Length from the base
of a dorsal spine above the pterygiophore to its distal tip along the
anterior edge \\
Pterygiophore & \texttt{lpt.sc} & Distance between the anterior to
posterior tips of the pterygiophore immediately preceding the 3rd dorsal
spine (when present) \\
\end{longtable}

\hypertarget{extant-specimen-data}{%
\subsection{Extant Specimen data}\label{extant-specimen-data}}

To span the gamut of stickleback diversity for our predictive imputation
model, we sampled modern stickleback from lakes containing generalist
stickleback populations (Hendry et al.
(\protect\hyperlink{ref-Hendryetal2009}{2009}); Bolnick
(\protect\hyperlink{ref-Bolnick2011}{2011})) and from lakes containing
benthic-limnetic species pairs (Baumgartner, Bell, and Weinberg
(\protect\hyperlink{ref-Baumgartneretal1988}{1988}); Schluter and
McPhail (\protect\hyperlink{ref-SchluterandMcPhail1992}{1992})). The
generalist specimens used here were collected by YES in 2013 and were
previously described in Stuart et al.
(\protect\hyperlink{ref-Stuartetal2017}{2017}) (Table S1). These samples
were fixed in formalin, then stained for bone with Alizarin Red in 2013.
Benthic and limnetic specimens were kindly loaned by D. Schluter and S.
Blain at the University of British Columbia. The Schluter lab collected
benthic and limnetic individuals from Enos Lake in 1988 and from Emily
Lake, Little Quarry Lake, Paxton Lake, and Priest Lake in 2018 (Table
S1). The Enos specimens had been fixed whole in formalin and stored in
40\% isopropanol. The specimens from the other lakes were initially
preserved whole in 95\% ethanol in the field before being gradually
transferred to water then formalin in the lab and ultimately stored in
40\% isopropanol. In 2019, we stained these specimens for bone using
Alizarin Red. We next replicated fossil data collection (Table 1) on
these extant specimens. Standard length as well as pelvic-spine length
on each side were measured with calipers. We used a dissection
microscope to count dorsal spines, pelvic spines, dorsal-fin rays, and
anal-fin rays. Right and left-side pelvic girdle lengths and
ectocoracoid lengths were measured from ventral photographs taken using
a Canon EOS Rebel T7 with a Tamron 16-300 mm MACRO lens mounted on a
leveled Kaiser RS1 copy stand. Specimens were held in place for ventral
photographs using a small tabletop vise with an attached scale bar.
Lateral X-rays were used to measure dorsal spine length, number of
pterygiophores anterior to the pterygiophore holding the third spine,
length of the pterygiophore just anterior to the third spine, cleithrum
length, and pre-maxilla ascending branch length. We also counted
vertebrae from the X-rays: abdominal vertebrae were counted anterior to
the first vertebra with a haemal spine contacting an anal fin
pterygiophore. Caudal vertebrae were posterior, including the first
vertebra with the haemal spine contacting the anal fin pterygiophore
(following Aguirre, Walker, and Gideon
(\protect\hyperlink{ref-Aguirreetal2014}{2014})). X-rays were taken with
an AXR Hot Shot X-ray Machine (Associated X-ray Corporation) at the
Field Museum of Natural History. Specimens were exposed at 35kV and 4mA.
Small fish were exposed for 7s, medium fish for 8s, and large fish for
10s. We developed the film and scanned individual images of each fish
using the B\&W Negatives setting on an Epson Perfection 4990 Photo
flatbed at 2400 dpi. Measurements from photographs and X-rays were taken
with FIJI (Schindelin et al.
(\protect\hyperlink{ref-Schindelinetal2012}{2012})) and its plugin
ObjectJ (\url{https://sils.fnwi.uva.nl/bcb/objectj/}).

We dissected individuals from the generalist populations (Table S1) to
determine sex from the gonads. Individuals from the species-pair lakes
(TableS1) were previously sexed by the Schluter lab, using either
dissection or a genotyping protocol (Whom, personal communication).

The extant data used here consists of a total of 367 specimens all with
known sex (Table S1). Of these, there are 202 and 165 female and male
specimens, respectively.

\hypertarget{outlier-analysis-and-size-correction.}{%
\subsection{Outlier analysis and size
correction.}\label{outlier-analysis-and-size-correction.}}

To check for outliers, we calculated within-group means and standard
deviations for each trait separately for K series fossil specimens
(pooled across samples) and for extant specimens (within generalist,
benthic, or limnetic categories). We noted trait values greater than 3.0
standard deviations from the mean as potential outliers. We deemed 3.0
s.d. to be a reasonable threshold for detecting errors without excluding
biologically relevant values. We checked whether these potential
outliers were a result of data entry and collection error and corrected
them if they were. We turned the remaining outlier trait values to NAs.
We size-corrected continuous traits only, as they varied with size,
unlike count traits that are fixed during early development. We
regressed each trait on standard length using a mixed-model regression,
pooling all specimens, following Stuart et al.
(\protect\hyperlink{ref-Stuartetal2017}{2017}). We appended the size
corrected data to the uncorrected trait data frame and used all of our
data (raw and size-corrected) to build the imputation model described
next.

\hypertarget{missing-data-imputation-including-fossil-sex}{%
\subsection{Missing data imputation, including fossil
sex}\label{missing-data-imputation-including-fossil-sex}}

Briefly, We used multiple imputation (Little and Rubin
(\protect\hyperlink{ref-little2002statistical}{2002})) to impute the sex
of the fossils. Let \(\boldsymbol{W}\) be an
\((n_{extant} + n_{fossil}) \times 1\) vector of the covariate sex of
the stickleback fish and \(\boldsymbol{Y}\) be an
\((n_{extant} + n_{fossil}) \times K\) matrix of the \(K\) phenotypes of
interest. Because the sex of the fossilized stickleback fish is
unobservable, we further define
\(\boldsymbol{W} = (\boldsymbol{W}_{extant}^T,\boldsymbol{W}_{fossil}^T)^T\)
where \(\boldsymbol{W}_{extant}\) and \(\boldsymbol{W}_{fossil}\) are
the \(n_{extant} \times 1\) and \(n_{fossil} \times 1\) vectors of the
observed extant sex and missing fossil sex, respectively. We imputeD
missing sex for the fossil data by sampling from the posterior
predictive distribution
\(P(\boldsymbol{W}_{fossil}|\boldsymbol{W}_{extant},\boldsymbol{Y})\)
using the multiple imputation by chained equations (MICE) algorithm
(Buuren and Groothuis-Oudshoorn (\protect\hyperlink{ref-MICE}{2011}))
with predictive mean matching, implemented in R (CITATION).
Traditionally, the choice for the number of completed dats sets is a
relatively small number such as \(M = 5\) or \(M = 10\). However, Zhou
and Reiter (\protect\hyperlink{ref-ZhouReiter2010}{2010}) recommend a
larger number of imputed data sets if users intend on performing
Bayesian analysis after imputation. Therefore, we imputed \(M = 100\)
complete datasets.

\textcolor{red}{presumably this also imputes missing data in the other traits? We should probably say that explicitly. -YS}
\textcolor{red}{AKHIL'S METHODS FOR VALIDATION OF THE IMPUTATION MODEL GOES HERE}
\textcolor{red}{THE POOLING DESCRIBED IN THE NEXT SENTENCE. WAS IT DONE, AND IF SO, HOW WAS IT USED IN DOWNSTREAM ANALYSES. MATT SAYS THAT HE RAN IS MODELS ON EACH IMPUTED DATA SET. -YS}
We then pooled all of the draws from the posterior distribution across
all the imputed data sets to estimate the posterior distributions of our
parameters of interest (i.e., the trait values), rather than using
Rubin's combining rules, following Zhou and Reiter
(\protect\hyperlink{ref-ZhouReiter2010}{2010}). We used these posterior
distributions in our Bayesian analysis.

\hypertarget{modeling-trait-means-through-time-for-males-and-females}{%
\subsection{Modeling trait means through time for males and
females}\label{modeling-trait-means-through-time-for-males-and-females}}

The goal of our analysis was to estimate fossil trait means through
time, by sex, while incorporating uncertainty in the imputed data,
especially uncertainty sex classification. We took a Bayesian approach
because imputation is the process of sampling from a posterior
predictive distribution of missing data, given observed data.
\textcolor{red}{THIS NEEDS MORE WORK TO JUSTIFY THE BAYESIAN APPROACH. -YS}

\textcolor{red}{DOES THE NEXT SENTENCE CONTRADICT THE PREVIOUS ONE? BELOW, WE SAY THAT WE USED EACH IMPUTED DATA SET. ABOVE, WE SAY THAT WE POOLED ACROSS IMPUTED DATA SETS.-YS}
To estimate fossil sample means through time, we first dropped the
extant individuals from our imputed data sets. Then, for a given imputed
dataset, we let \(W_{ij}\) be the imputed sex and
\(\boldsymbol{Y}_{ij}\) be the \(K \times 1\) vector of phenotypes for
stickleback fossil \(j\) at time \(t_i\) where \(i = 1, \ldots, T\) and
\(j = 1,\ldots,n_{t}\). In addition, we denote \(Y_{K,ij}\), the last
variable in \(\boldsymbol{Y}_{ij}\), to be the standard length of each
fish individual, a measure of body size. \begin{align}
{Y}_{K,ij} & \overset{iid}{\sim}\left\{\begin{array}{lll} \mathcal{N}(\mu_{K,ft_i},&\sigma_{K}^2), & W_{ij} = \text{Female} \\ \mathcal{N}(\mu_{K,mt_i},&\sigma_{K}^2), & W_{ij} = \text{Male} \end{array}\right..
\label{eq:stl}
\end{align}

It is reasonable to assume that the other continuous traits will have
some correlation with standard length \textbf{(CITATION)}. We account
for this by adding an additional parameter, \(\gamma_k\), If
\(Y_{k,ij}\) is a continuous trait, then \begin{align}
{Y}_{k,ij} & \overset{iid}{\sim}\left\{\begin{array}{llll} \mathcal{N}(\mu_{k,ft_i} + \gamma_kY_{K,ij},\sigma_k^2), & W_{ij} = \text{Female} \\ \mathcal{N}(\mu_{k,mt_i} + \gamma_kY_{K,ij},\sigma_k^2), & W_{ij} = \text{Male} \end{array}\right..
\label{eq:cont}
\end{align}

If \(Y_{k,ij}\) is a discrete trait, the conventional modelling method
is to fit a Poisson distribution. However, because the empirical fossil
data violate the assumption that the Poisson variance is equal to its
mean, we fit discrete traits to a generalized Poisson model as defined
in
(\protect\hyperlink{ref-GeneralizedPoisson}{\textbf{GeneralizedPoisson?}}).
Specifically, if \(X \sim GP(\lambda,\alpha)\), then \begin{align}
P(X = x) = \left\{\begin{array}{cc} \frac{(1 - \alpha)\lambda[(1 - \alpha)\lambda + \alpha x]^{x - 1} \exp\left\{-((1 - \alpha)\lambda  + \alpha x)\right\}}{x!} & (1 - \alpha)\lambda  + \alpha x \geq 0  \\ 0 & (1 - \alpha)\lambda  + \alpha x < 0 \end{array}\right.,
\label{eq:GP_pmf}
\end{align} where GP denotes a generalized Poisson distribution with
\(E(X) = \lambda\) and \(Var(X) = \frac{\lambda}{(1 - \alpha)^2}\).

In addition, we allowed that two discrete traits, abdominal vertebrae
number (mav) and caudal vertebrae number (mcv) could also with standard
length (\textbf{CITATION}). This is because these are vertebral counts
that run in parallel with the standard length measurement, along the
anterior-posterior access of the fish; these vertebrae thus contribute
to the length of the fish directly, and may evolve in a correlated
fashion. If \(Y_{k,ij}\) is one of the above traits, then \begin{align}
{Y}_{k,ij} & \sim \left\{\begin{array}{ll} GP(\exp\{\mu_{k,ft_i} + \gamma_kY_{K,ij}\},\alpha_k), & W_{ij} = \text{Female} \\ GP(\exp\{\mu_{k,mt_i} + \gamma_kY_{K,ij}\},\alpha_k), & W_{ij} = \text{Male} \end{array}\right..
\label{eq:disc_corr}
\end{align}

For the other discrete traits, we set \(\gamma_k = 0\) because we expect
that trait counts are set during early ontogeny, and should not change
with size in adult fish. Thus, \begin{align}
{Y}_{k,ij} & \sim \left\{\begin{array}{ll} GP(\exp\{\mu_{k,ft_i}\},\alpha_k), & W_{ij} = \text{Female} \\ GP(\exp\{\mu_{k,mt_i}\},\alpha_k), & W_{ij} = \text{Male} \end{array}\right..
\label{eq:disc_ind}
\end{align} We point out that, for the discrete traits, the means are
represented by \(\exp\{\mu_{k,ft_i}\}\) and \(\exp\{\mu_{k,mt_i}\}\) for
ease of modeling.

In the above model descriptions, \(\mu_{k,ft_i}\) and \(\mu_{k,mt_i}\)
model the time-\(t_i\), sample-specific trait mean \(k\) for females and
males, respectively. To account for the possibility that there is a time
-dependent trend in \(\mu_{k,ft_i}\) and \(\mu_{k,mt_i}\), we further
set \begin{align}
\mu_{k,gt_i} = \beta_{0,kg} + \beta_{1,kg}t_i + u_{k,gt_i},
\label{eq:mu}
\end{align} for \(g \in \{f,m\}\), where \(\beta_{0,kg}\) and
\(\beta_{1,kg}\) are regression parameters of phenotype \(Y_k\) for each
sex, and \(u_{k,gt_i}\) is the corresponding residual. To account for
potential correlations between the residuals for a given trait \(k\) and
sex \(g\), we fit an Ornstein-Uhlenback (OU) process (Uhlenbeck and
Ornstein (\protect\hyperlink{ref-OUProcess}{1930})). More specifically,
we defined \(du_{k,gt} = u_{k,g(t + dt)} - u_{k,gt}\), the change in
\(u_k,gt\) for a given trait \(k\) and sex \(g\) over a miniscule time
period \(dt\). The OU process is defined as \begin{align}
du_{k,gt} = -\kappa_k u_{k,gt} dt + \tau_k dW_t,
\label{eq:OU_cont}
\end{align} where \(\kappa_k\) is a parameter associated with the
correlation between \(u_{k,gt}\) and \(u_{k,g(t+dt)}\), \(\tau_k\) is
the standard deviation of the OU process, and \(W_t\) is a standard
Brownian motion. As shown in Uhlenbeck and Ornstein
(\protect\hyperlink{ref-OUProcess}{1930}), the closed form solution for
the stochastic differential equation in (\ref{eq:OU_cont}) is
\begin{align}
u_{k,gt_i} \overset{iid}{\sim}\mathcal{N}\left(u_{k,gt_{i-1}}\exp\{-\kappa_k(t_{i} - t_{i-1})\} , \frac{\tau_k^2(1 - \exp\{-2\kappa_k(t_{i} - t_{i-1}))\}}{2\kappa_k}\right)
\label{eq:OU_sol}
\end{align} for \(i = 2,\cdots,T\). In a traditional OU process, the
initial value \(u_{k,gt_1}\) is assumed to be a (potentially unknown)
constant. This unknown constant can therefore be considered a latent
variable, further justifying our Bayesian approach, which better
incorporates prior information and uncertainty from missing data
imputation in the estimate of the variable.

We discuss the procedure for estimating this value below.
\textcolor{red}{DID THIS GET DISCUSSED?-YS}

Priors: To aid in the sampling procedure for the discrete phenotypes, we
first had to transform the parameters of the generalized Poisson
distribution onto the real number line in order to ensure our algorithm
can properly sample from the posterior distribution of interest using
STAN software. This was done by declaring
\(\phi_k = \log\left(\frac{\alpha_k - \max_{i,j}(-\lambda_{k,ij}/y_{k,ij})}{1 - \alpha_k}\right)\)
where
\(\lambda_{k,ij} = \left\{\begin{array}{cc} \exp(\mu_{k,ft_i} + \gamma_k Y_{K,ij}) & W_{ij} = \text{Female} \\ \exp(\mu_{k,mt_i} + \gamma_k Y_{K,ij}) & W_{ij} = \text{Male} \end{array}\right.\).

We chose the following priors, designing them to be weakly informative
such that the posterior distributions could be influenced by the data,
if warranted. For \(k = 1,\cdots,K\),

\begin{align}
u_{k,gt_1} & \overset{iid}{\sim}\mathcal{N}(0,\tau_{0,k}) \nonumber \\
\sigma_k & \overset{iid}{\sim}\mathcal{N}(0,10)I_{\{\sigma > 0\}} \nonumber \\
\tau_k & \overset{iid}{\sim}\mathcal{N}(0,10)I_{\{\tau > 0\}} \nonumber \\
\tau_{0,k} & \overset{iid}{\sim}\mathcal{N}(0,20)I_{\{\tau > 0\}} \nonumber \\
\kappa_k & \overset{iid}{\sim}\mathcal{N}(0,1)I_{\{\kappa_g > 0\}} \nonumber \\
\gamma_{k} & \overset{iid}{\sim}\mathcal{N}(0,5) \nonumber \\
\beta_{0,kg} & \overset{iid}{\sim}\mathcal{N}(0,100) \nonumber \\
\beta_{1,kg} & \overset{iid}{\sim}\mathcal{N}(0,3) \nonumber \\
\label{eq:priors}
\end{align}

We also note that, for the discrete phenotypes in equations
(\ref{eq:disc_corr}) and (\ref{eq:disc_ind}), there is no \(\sigma_k\),
and it is not sampled, and for the continuous phenotypes in equation
(\ref{eq:cont}), there is no \(\phi_k\) and it is not sampled.

We follow the method outlined in
(\protect\hyperlink{ref-ZhouandReiter}{\textbf{ZhouandReiter?}}) to
obtain \(R\) draws from the posterior distribution of the outlined model
parameters given our observed data and imputed values. More
specifically, for the \(j^{th}\) imputed data set where
\(j = 1,\cdots,M\), we obtain \(\frac{R}{M}\) draws from the posterior
distribution using an MCMC algorithm with a sufficient burnin period,
and then we combine these \(M\) MCMC draws to say that we have a total
of \(R\) draws from the overall posterior distribution of interest,
which can then be used to assess the presence of sexual dimorphism for a
given trait.

All models were built using Team
(\protect\hyperlink{ref-R2022language}{2022}) and (\textbf STAN)

After running these models, we have posterior distributions of all the
model parameters of interest. From these, we can obtain samples from the
posterior distribution of mean trait values for each trait for males and
females, using equation 6. We use these values for the following
analyses.

\hypertarget{quantifying-sexual-dimorphism-and-testing-for-increased-sexual-dimorphism-through-time.}{%
\subsection{Quantifying sexual dimorphism, and testing for increased
sexual dimorphism through
time.}\label{quantifying-sexual-dimorphism-and-testing-for-increased-sexual-dimorphism-through-time.}}

We identify sexual dimorphism first by calculating the difference in
mean trait size for males and females at each time for phenotype \(k\)
(\(\mu_{k,m,t_i} - \mu_{k,m,t_i}\)) for \(i = 1,\cdots,T\) where, for
the continuous phenotypes, the mean trait sizes for each sex
\(\mu_{k,m,t_i}\) and \(\mu_{k,f,t_i}\) are calculated using
(\ref{eq:mu}). For the discrete phenotypes, we note that in
(\ref{eq:disc_corr}) and (\ref{eq:disc_ind}) the mean trait size is
\(\lambda_{k,m,t_i} = \exp(\mu_{k,m,t_i})\) and
\(\lambda_{k,f,t_i} = \exp(\mu_{k,f,t_i})\) for males and females,
respectively, which is the value we use in our analysis. For simplicity,
we will use \(\mu_{k,m,t_i}\) and \(\mu_{k,f,t_i}\) as the notation for
mean size for both discrete and continuous traits for the rest of this
discussion. From our MCMC algorithm, we obtain \(M\) samples from the
posterior distribution of \(\mu_{k,m,t_i} - \mu_{k,f,t_i}\) to use in
our analysis.

Recall that we predicted that sexual dimorphism should increase through
time in Lineage II, following ecological release from predators.
However, this change in dimorphism make be increasing at a non-constant
rate, as pointed out by the inclusion of the OU process. We assess for
dimorphism by calculating a Kendall's \(\tau\) coefficient (KENDALL
(\protect\hyperlink{ref-KendallsTau}{1938})). Kendall's \(\tau\)
assesses the ordinal association between a set of bivariate
observations, where \(\tau > 0\) indicates concordance between the
observations (i.e., as \(x\) increases, \(y\) also tends to increase.)

For our analysis, we calculate
\(\boldsymbol{\tau}_k = \{\tau_{k,r}: r = 1,\cdots,R\}\) for trait \(k\)
where \(\tau_r\) is the estimate of the Kendall's \(\tau\) correlation
for \(\mu_{k,m,t_i} - \mu_{k,f,t_i}\) versus time \(t_i\) for the
\(r^{th}\) MCMC sample from the posterior distribution. The posterior
probability of sexual dimorphism for trait \(k\) can be estimated as the
proportion of values in \(\boldsymbol{\tau}_k\) that are greater than 0
(\(\frac{1}{R} \sum_{r=1}^R I(\tau_{k,r} > 0)\)). We can also assess the
strength of the dimorphism by calculating the proportion of posterior
samples that are greater than a prespecified threshold
\textcolor{red}{(0.5??, idk what this could be.) -MS}

\hypertarget{sec:results}{%
\section{Results}\label{sec:results}}

\textcolor{red}{WE NEED RESULTS SECTION DESCRIBING THE RESULTS OF AKHIL'S MULTIPLE IMPUTATION, BOTH THE TESTS OF HOW WELL THE MODEL WORKED CLASSIFYING MODEL SAMPLES, AND SUMMARY STATISTICS OF FOSSIL IMPUTATION RESULTS -YS}

\hypertarget{sexual-dimorphism}{%
\subsection{Sexual Dimorphism}\label{sexual-dimorphism}}

Table \ref{tab:prob0-table} shows the posterior probability of the
differences in the means between male and female specimens at each time
point. Probabilities near 0.5 indicate very little difference in the
means whereas probabilities far from 0.5 indicate sexual dimorphism with
values of 0 and 1 indicating larger mean values for females and males,
respectively. These results are shown graphically in figure
\ref{fig:prob0-plot}.

Figure \ref{fig:mu-diff-plot} shows the posterior mean difference for
males versus females (values above 0 indicate means that were greater
for males versus females). The full posterior distributions over time
for all traits are shown in the appendix.

Traits such as standard length (stl), anal fin ray (maf), dorsal fin ray
(mdf), premaxilla (pmx), and dorsal spine 1 (ds1) demonstrate constant
sexual dimorphism with all but the last having learger mean values in
male sepcimen. Other traits such as abdominal vertebra (mav), caudal
vertebra (mcv), and pelvic spine length (lps) don't appear to have any
discernable sexual dimorphism in the means. Dorsal spine 3 (ds3)
exhibits sexual dimorphism that changes over time. The mean number of
dorsal spine 3 begins larger for females for the earliest observed time
periods. Over time, however, this switches and the male specimens
exhibit larger mean values for the latest observed time periods.

\textcolor{red} {
Yoel's interpretation of the three paragraphs above. Male and female G. doryssus differ for some traits, indicating detectable sexual dimorphism. This is evidenced by traits with substantial differences in male and female posterior means (i.e., deviation from 0 in Figure \ref{fig:mu-diff-plot}), supported by posterior probabilities that deviate from 0.5 (Figure \ref{fig:prob0-plot}). For example, males are larger than females by 3-8mm on average through time (stl panel in Figure \ref{fig:mu-diff-plot}), with a posterior probability for that mean difference above 70 percent through time (Figure \ref{fig:prob0-plot}). In contrast, the number of abdominal vertebrae (mav) fluctuates about 0 (Figure \ref{fig:mu-diff-plot}), with posterior probabilities fluctuating about 50 percent through time (Figure \ref{fig:prob0-plot}). Table \ref{tab:prob0-table} in the Supplementary Material summarizes the posterior probabilities that male means differ from female means by trait and sample.
}

\textcolor{red}{can we move mu-diff-plot and prob0-plot here?}

\hypertarget{changes-over-time}{%
\subsection{Changes over time}\label{changes-over-time}}

Table \ref{tab:auc-table} shows the probability that the posterior
distribution of the differences in the mean is greater than the
posterior distribution of the differences at the first time point.
Values near 0.5 indicate very little difference between the
distributions whereas values closer to 0 and 1 indicate changes in the
distributions with values of 1 meaning the difference in means has
shifted towards males having a larger mean and values of 0 indicating
the difference in means has shifted towards larger means in females.
These results are shown graphically in figure \ref{fig:AUC-plot}.

The amount of sexual dimporphism over time does not change substantially
for a majority of the trait observed here (i.e.~cle, ds1, ds2, ect, lps,
lpt, mav, mcv, pmx, stl). Traits that trend away from the mean
difference at the first observed time point include dorsal spine 3
(ds3), anal fin ray (maf), and dorsal fin ray (mdf).

The posterior mean of the mean number of dorsal spine rays at the first
observed time point is 2.29 for males and for females it is 2.45. At the
last observed time point, the average number for males and female both
dropped to 1.265 and 1.185 for males and females, respectively. This
resulted in a mean number of dorsal spine 3 lost by males of 1.027 and
for females it was 1.264. This resulted in a change of the mean
difference from time point 1 to time point 18 of 0.236. Basically, males
and females both lowered the mean number of dorsal spine 3, but females
lowered their mean more than the males.

We also observed a change in the differences between males and females
for the number of bones int he anal fin ray (maf) and the number of
bones in the dorsal fin ray. The difference in the means of males versus
females for anal fin ray bones changed by 0.152 over the course of the
observed time periods. This change was driven mostly by the change in
females. At the first observed time period the had on average 8.291
bones dropping to 8.170 over the 18 times periods observed. Whereas for
males the change in means rose only from 8.518 to 8.549, about 0.0316
bones on average.

A similar change in the difference of the means over time was observed
for the number of bones in the dorsal fin (mdf). However, unlike maf,
this change in sexual dimporhism was driven entirely by the males. The
mean change for females over the 18 observed time periods was only
0.0055 with the average number of dorsal fin bones for females starting
at 9.287 and the final observed mean of 9.281 resulting in almost no
change in the mean. For males, the average number of dorsal fin bones at
the first time point was about 9.568 dropping to 9.734 over the course
of the observed time periods for a gain of 0.166 bones on average.

\textcolor{red} {
Yoel's interpretation of the five paragraphs above. 
Figure \ref{fig:AUC-plot} shows the posterior probability that the mean male-female difference at time t is different from the male-female difference at time zero. In other words, Figure \ref{fig:AUC-plot} plots whether the sexual dimorphism changes in subsequent samples, always relative to the first sample. The amount of sexual dimporphism over time does not change substantially for a majority of the traits (i.e. cle, ds1, ds2, ect, lps, lpt, mav, mcv, pmx, stl; Figure \ref{fig:AUC-plot}). Traits that change in dimorphism are anal fin ray (maf), and dorsal fin ray (mdf), and dorsal spine 3 (ds3).}

\textcolor{red} {
For example, females at time zero averaged 8.29 rays in the anal fin, dropping to 8.17 rays in the final sample. Males averaged 8.52 anal fin rays in the first sample and 8.55 in the last sample. Thus, the difference between males and females started at 0.23 and ended at 0.38, an increase in sexual dimorphism. In contrast, the sign on dimorphism flipped for ds3. The posterior mean length of the third dorsal spine is 2.29mm for males and 2.45mm for females. By the end of the sequence, mean ds3 for males was 1.27 for males and 1.19 for females. Both sexes evolved reduction of ds3, but females lowered their mean more than the males. This resulted in a flip in dimorphism, from larger females to larger males, though both sexes evolved trait reduction. BUT DIMORPHISM SHRUNK THEN, AS THE DIFFERENCE BETWEEN MALES AND FEMALES IS NOW SMALLER??
}

\begin{longtable}[]{@{}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 26\tabcolsep) * \real{0.0602}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 26\tabcolsep) * \real{0.0723}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 26\tabcolsep) * \real{0.0723}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 26\tabcolsep) * \real{0.0723}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 26\tabcolsep) * \real{0.0723}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 26\tabcolsep) * \real{0.0723}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 26\tabcolsep) * \real{0.0723}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 26\tabcolsep) * \real{0.0723}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 26\tabcolsep) * \real{0.0723}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 26\tabcolsep) * \real{0.0723}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 26\tabcolsep) * \real{0.0723}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 26\tabcolsep) * \real{0.0723}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 26\tabcolsep) * \real{0.0723}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 26\tabcolsep) * \real{0.0723}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedleft
time
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
cle
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
ds1
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
ds2
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
ds3
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
ect
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
lps
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
lpt
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
maf
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
mav
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
mcv
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
mdf
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
pmx
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
stl
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1 & 0.913 & 0.215 & 0.359 & 0.203 & 0.865 & 0.428 & 0.853 & 0.944 &
0.391 & 0.507 & 0.959 & 0.948 & 0.883 \\
2 & 0.942 & 0.212 & 0.325 & 0.180 & 0.891 & 0.404 & 0.853 & 0.948 &
0.520 & 0.520 & 0.963 & 0.967 & 0.908 \\
3 & 0.843 & 0.207 & 0.249 & 0.153 & 0.770 & 0.279 & 0.803 & 0.965 &
0.289 & 0.525 & 0.977 & 0.912 & 0.777 \\
4 & 0.890 & 0.210 & 0.249 & 0.201 & 0.830 & 0.372 & 0.741 & 0.977 &
0.435 & 0.589 & 0.986 & 0.935 & 0.831 \\
5 & 0.954 & 0.211 & 0.211 & 0.195 & 0.909 & 0.377 & 0.756 & 0.985 &
0.533 & 0.622 & 0.992 & 0.975 & 0.921 \\
6 & 0.885 & 0.211 & 0.134 & 0.171 & 0.820 & 0.291 & 0.686 & 0.991 &
0.645 & 0.456 & 0.996 & 0.949 & 0.811 \\
7 & 0.812 & 0.211 & 0.124 & 0.239 & 0.760 & 0.340 & 0.617 & 0.994 &
0.224 & 0.717 & 0.999 & 0.879 & 0.747 \\
8 & 0.928 & 0.210 & 0.135 & 0.258 & 0.885 & 0.387 & 0.706 & 0.994 &
0.342 & 0.730 & 1.000 & 0.969 & 0.872 \\
9 & 0.899 & 0.213 & 0.145 & 0.304 & 0.857 & 0.406 & 0.715 & 0.995 &
0.521 & 0.756 & 1.000 & 0.940 & 0.845 \\
10 & 0.954 & 0.217 & 0.235 & 0.555 & 0.940 & 0.571 & 0.810 & 0.997 &
0.454 & 0.719 & 1.000 & 0.976 & 0.935 \\
11 & 0.913 & 0.211 & 0.162 & 0.476 & 0.875 & 0.388 & 0.703 & 0.997 &
0.363 & 0.657 & 1.000 & 0.957 & 0.864 \\
12 & 0.949 & 0.221 & 0.262 & 0.726 & 0.929 & 0.585 & 0.803 & 0.997 &
0.462 & 0.595 & 1.000 & 0.968 & 0.926 \\
13 & 0.904 & 0.220 & 0.260 & 0.789 & 0.885 & 0.598 & 0.795 & 0.998 &
0.470 & 0.370 & 1.000 & 0.929 & 0.872 \\
14 & 0.925 & 0.220 & 0.280 & 0.883 & 0.906 & 0.659 & 0.854 & 0.997 &
0.526 & 0.472 & 1.000 & 0.958 & 0.898 \\
15 & 0.880 & 0.225 & 0.283 & 0.839 & 0.849 & 0.617 & 0.796 & 0.997 &
0.490 & 0.398 & 1.000 & 0.919 & 0.834 \\
16 & 0.953 & 0.216 & 0.276 & 0.931 & 0.931 & 0.614 & 0.831 & 0.997 &
0.414 & 0.541 & 1.000 & 0.975 & 0.925 \\
17 & 0.803 & 0.219 & 0.232 & 0.832 & 0.765 & 0.383 & 0.655 & 0.996 &
0.569 & 0.758 & 1.000 & 0.854 & 0.751 \\
18 & 0.780 & 0.212 & 0.225 & 0.817 & 0.726 & 0.297 & 0.558 & 0.995 &
0.705 & 0.722 & 0.999 & 0.859 & 0.717 \\
\end{longtable}

\begin{longtable}[]{@{}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 26\tabcolsep) * \real{0.0602}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 26\tabcolsep) * \real{0.0723}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 26\tabcolsep) * \real{0.0723}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 26\tabcolsep) * \real{0.0723}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 26\tabcolsep) * \real{0.0723}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 26\tabcolsep) * \real{0.0723}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 26\tabcolsep) * \real{0.0723}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 26\tabcolsep) * \real{0.0723}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 26\tabcolsep) * \real{0.0723}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 26\tabcolsep) * \real{0.0723}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 26\tabcolsep) * \real{0.0723}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 26\tabcolsep) * \real{0.0723}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 26\tabcolsep) * \real{0.0723}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 26\tabcolsep) * \real{0.0723}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedleft
time
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
cle
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
ds1
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
ds2
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
ds3
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
ect
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
lps
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
lpt
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
maf
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
mav
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
mcv
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
mdf
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
pmx
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
stl
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1 & 0.500 & 0.500 & 0.500 & 0.500 & 0.500 & 0.500 & 0.500 & 0.500 &
0.500 & 0.500 & 0.500 & 0.500 & 0.500 \\
2 & 0.514 & 0.504 & 0.509 & 0.495 & 0.512 & 0.508 & 0.517 & 0.495 &
0.396 & 0.490 & 0.495 & 0.511 & 0.513 \\
3 & 0.636 & 0.514 & 0.559 & 0.516 & 0.628 & 0.601 & 0.602 & 0.475 &
0.523 & 0.480 & 0.476 & 0.626 & 0.641 \\
4 & 0.563 & 0.509 & 0.544 & 0.459 & 0.558 & 0.539 & 0.672 & 0.456 &
0.436 & 0.450 & 0.458 & 0.556 & 0.562 \\
5 & 0.543 & 0.507 & 0.551 & 0.417 & 0.539 & 0.522 & 0.687 & 0.436 &
0.389 & 0.425 & 0.439 & 0.537 & 0.544 \\
6 & 0.618 & 0.512 & 0.589 & 0.415 & 0.613 & 0.579 & 0.716 & 0.415 &
0.317 & 0.522 & 0.419 & 0.612 & 0.620 \\
7 & 0.591 & 0.510 & 0.601 & 0.364 & 0.589 & 0.564 & 0.712 & 0.379 &
0.546 & 0.382 & 0.383 & 0.588 & 0.592 \\
8 & 0.539 & 0.507 & 0.579 & 0.324 & 0.536 & 0.514 & 0.683 & 0.373 &
0.478 & 0.368 & 0.378 & 0.537 & 0.540 \\
9 & 0.524 & 0.506 & 0.586 & 0.299 & 0.521 & 0.510 & 0.636 & 0.355 &
0.408 & 0.340 & 0.359 & 0.523 & 0.522 \\
10 & 0.379 & 0.497 & 0.544 & 0.203 & 0.377 & 0.400 & 0.578 & 0.321 &
0.438 & 0.359 & 0.324 & 0.385 & 0.373 \\
11 & 0.531 & 0.508 & 0.605 & 0.226 & 0.528 & 0.517 & 0.685 & 0.301 &
0.528 & 0.403 & 0.303 & 0.530 & 0.533 \\
12 & 0.363 & 0.492 & 0.547 & 0.150 & 0.360 & 0.386 & 0.566 & 0.284 &
0.439 & 0.436 & 0.284 & 0.369 & 0.358 \\
13 & 0.373 & 0.495 & 0.555 & 0.133 & 0.372 & 0.386 & 0.561 & 0.269 &
0.432 & 0.575 & 0.268 & 0.378 & 0.371 \\
14 & 0.326 & 0.496 & 0.548 & 0.102 & 0.326 & 0.349 & 0.514 & 0.255 &
0.399 & 0.514 & 0.253 & 0.334 & 0.323 \\
15 & 0.365 & 0.492 & 0.552 & 0.107 & 0.365 & 0.373 & 0.502 & 0.250 &
0.433 & 0.554 & 0.247 & 0.368 & 0.367 \\
16 & 0.361 & 0.497 & 0.574 & 0.084 & 0.360 & 0.381 & 0.570 & 0.234 &
0.462 & 0.463 & 0.229 & 0.369 & 0.356 \\
17 & 0.561 & 0.504 & 0.629 & 0.116 & 0.559 & 0.542 & 0.669 & 0.226 &
0.364 & 0.313 & 0.219 & 0.559 & 0.565 \\
18 & 0.625 & 0.512 & 0.648 & 0.121 & 0.617 & 0.596 & 0.735 & 0.220 &
0.281 & 0.356 & 0.212 & 0.622 & 0.627 \\
\end{longtable}

posterior mean difference for males versus females (values above 0
indicate means that were greater for males versus females)

\begin{figure}

{\centering \includegraphics{paper_files/figure-latex/mu-diff-plot-1}

}

\caption{Posterior mean difference between males and females, in UNITS?. Values above 0 indicate larger male means}\label{fig:mu-diff-plot}
\end{figure}

\begin{figure}

{\centering \includegraphics{paper_files/figure-latex/prob0-plot-1}

}

\caption{Posterior probabilities that the male posterior mean is larger than the female posterior mean}\label{fig:prob0-plot}
\end{figure}

\begin{figure}

{\centering \includegraphics{paper_files/figure-latex/AUC-plot-1}

}

\caption{The Posterior Probability that male-female differences in subsequent samples are different from male-female differenceds in the initial sample. Values near 0.5 indicate little difference between the distributions. Values closer to 0 indicate that the difference in means has shifted towards larger means in females. Values closer to 1 indicate that the difference in means has shifted toward larger means in males.}\label{fig:AUC-plot}
\end{figure}

\hypertarget{sec:conclusions}{%
\section{Discussion, Future work and
conclusions}\label{sec:conclusions}}

We predicted that release from predators would result in niche expansion
and increased sexual dimorphism, based on several studies of modern
stickleback. For example, in lakes where sculpin competitors are absent
and stickleback (Roesti et al.~2023) See Spoljaric and Reimchen 2008,
page 512 right column for references and discussion of differences
between benthic males and limnetic females. Male stickleback are benthic
and littoral (Wootton 1976)\ldots. Reimchen papers in general good for
this section. Moreover, tooth wear data from the lineage II sequence
suggest that individuals in this lineage began eating more planktonic
prey over time, expanding toward an open-water niche from the benthic,
bottom-feeding niche they started with (Purnell et al.~2007). That this
expansion into open water by lineage II coincided with armor reduction
is further indicative of a limnetic system with fewer salmonid
piscivores (Schluter and McPhail
(\protect\hyperlink{ref-SchluterandMcPhail1992}{1992}); Vamosi and
Schluter (\protect\hyperlink{ref-VamosiandSchluter2004}{2004}); Roesti
et al. (\protect\hyperlink{ref-Roestietal2023}{2023})).

\hypertarget{acknowledgements}{%
\section*{Acknowledgements}\label{acknowledgements}}
\addcontentsline{toc}{section}{Acknowledgements}

We thank O. Abughoush, S. Blain, A. Chaudhary, M. Islam, F. Joaquin, C.
Lawson-Weinert, R. Sullivan, J. Tien, M.P. Travis, and W. Shim for help
with data collection. We thank D. Schluter and S. Blain for loaning
specimens and sharing data. We thank K. Swagel and C. McMahan of the
Field Museum for assistance with specimen x-rays. This research was
supported by NSF grants BSR-8111013, EAR-9870337, and DEB-0322818, the
Center for Field Research (Earthwatch), and the National Geographic
Society (2869-84) to MAB. It was also supported by NSF grants
DEB-1456462 and EAR-2145830 to YES. And NSF DMS-2015374 (GJM)

\hypertarget{supplementary-material}{%
\section*{Supplementary Material}\label{supplementary-material}}
\addcontentsline{toc}{section}{Supplementary Material}

All code for reproducing the analyses in this paper is publicly
available at \url{https://github.com/Akhil-Ghosh/SticklebackProject}

DIC results here.

\hypertarget{figures}{%
\subsection{Figures}\label{figures}}

\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{../Figures/cle/mu_diff}

}

\caption{cle}\label{fig:unnamed-chunk-7}
\end{figure}
\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{../Figures/ds1/mu_diff}

}

\caption{ds1}\label{fig:unnamed-chunk-8}
\end{figure}
\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{../Figures/ds2/mu_diff}

}

\caption{ds2}\label{fig:unnamed-chunk-9}
\end{figure}
\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{../Figures/ds3/mu_diff}

}

\caption{ds3}\label{fig:unnamed-chunk-10}
\end{figure}
\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{../Figures/ect/mu_diff}

}

\caption{ect}\label{fig:unnamed-chunk-11}
\end{figure}
\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{../Figures/lps/mu_diff}

}

\caption{lps}\label{fig:unnamed-chunk-12}
\end{figure}
\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{../Figures/lpt/mu_diff}

}

\caption{lpt}\label{fig:unnamed-chunk-13}
\end{figure}
\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{../Figures/maf/lambda_diff}

}

\caption{maf}\label{fig:unnamed-chunk-14}
\end{figure}
\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{../Figures/mav/lambda_diff}

}

\caption{mav}\label{fig:unnamed-chunk-15}
\end{figure}
\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{../Figures/mcv/lambda_diff}

}

\caption{mcv}\label{fig:unnamed-chunk-16}
\end{figure}
\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{../Figures/mdf/lambda_diff}

}

\caption{mdf}\label{fig:unnamed-chunk-17}
\end{figure}
\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{../Figures/mds/lambda_diff}

}

\caption{mds}\label{fig:unnamed-chunk-18}
\end{figure}
\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{../Figures/mpt/lambda_diff}

}

\caption{mpt}\label{fig:unnamed-chunk-19}
\end{figure}
\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{../Figures/pmx/mu_diff}

}

\caption{pmx}\label{fig:unnamed-chunk-20}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=0.9\linewidth]{../Figures/stl/mu_diff}

}

\caption{stl}\label{fig:unnamed-chunk-21}
\end{figure}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Aguirreetal2014}{}}%
Aguirre, W E, K Walker, and S Gideon. 2014. {``Tinkering with the Axial
Skeleton: Vertebral Number Variation in Ecologically Divergent
Threespine Stickleback Populations.''} \emph{Biological Journal of the
Linnean Society} 113 (1): 204--19.

\leavevmode\vadjust pre{\hypertarget{ref-Baumgartneretal1988}{}}%
Baumgartner, J V, M A Bell, and P H Weinberg. 1988. {``Body Form
Differences Between the Enos Lake Species Pair of Threespine
Sticklebacks (Gasterosteus Aculeatus Complex).''} \emph{Canadian Journal
of Zoology} 66 (2): 467--74.

\leavevmode\vadjust pre{\hypertarget{ref-Bell2009}{}}%
Bell, M A. 2009. {``Implications of a Fossil Stickleback Assemblage for
Darwinian Gradualism.''} \emph{Journal of Fish Biology} 75 (8):
1977--99.

\leavevmode\vadjust pre{\hypertarget{ref-Belletal1985}{}}%
Bell, M A, J V Baumgartner, and E C Oslon. 1985. {``Patterns of Temporal
Change in Single Morphological Characters of a Miocene Stickleback
Fish.''} \emph{Paleobiology} 11: 258--71.

\leavevmode\vadjust pre{\hypertarget{ref-Belletal1993}{}}%
Bell, M A, G Orti, J A Walker, and J P Koenings. 1993. {``Evolution of
Pelvic Reduction in Threespine Sticklebacks: A Test of Competing
Hypotheses.''} \emph{Evolution} 47: 906--14.

\leavevmode\vadjust pre{\hypertarget{ref-Belletal2006}{}}%
Bell, M A, M P Travis, and D M Blouw. 2006. {``Inferring Natural
Selection in a Fossil Threespine Stickleback.''} \emph{Paleobiology} 32
(4): 562--77.

\leavevmode\vadjust pre{\hypertarget{ref-Blain2022}{}}%
Blain, S A. 2022. \emph{Evolutionary Outcomes of Interactions Among
Phenotypes in Post-Glacial Lakes}. University of British Columbia,
Canada.

\leavevmode\vadjust pre{\hypertarget{ref-Bolnick2011}{}}%
Bolnick, D I. 2011. {``Sympatric Speciation in Threespine Stickleback:
Why Not?''} \emph{International Journal of Ecology} 2011: 1--15.

\leavevmode\vadjust pre{\hypertarget{ref-BolnickandDoebeli2003}{}}%
Bolnick, D I, and M Doebeli. 2003. {``Sexual Dimorphism and Adaptive
Speciation: Two Sides of the Same Ecological Coin.''} \emph{Evolution}
57 (11): 2433--49.

\leavevmode\vadjust pre{\hypertarget{ref-BolnickandLau2008}{}}%
Bolnick, D I, and O L Lau. 2008. {``Predictable Patterns of Disruptive
Selection in Stickleback in Postglacial Lakes.''} \emph{The American
Naturalist} 172 (1): 1--11.

\leavevmode\vadjust pre{\hypertarget{ref-Bowne1994}{}}%
Bowne, P S. 1994. {``Systematics and Morphology of the
Gasterosteiformes.''} In \emph{The Evolutionary Biology of the
Threespine Stickleback}, edited by M A Bell and S A Foster. Oxford, UK:
Oxford University Press.

\leavevmode\vadjust pre{\hypertarget{ref-Butleretal2007}{}}%
Butler, Sawyer, M A, and J B Losos. 2007. {``Sexual Dimorphism and
Adaptive Radiation in Anolis Lizards.''} \emph{Nature} 447 (7141):
202--5.

\leavevmode\vadjust pre{\hypertarget{ref-MICE}{}}%
Buuren, Stef van, and Karin Groothuis-Oudshoorn. 2011. {``Mice:
Multivariate Imputation by Chained Equations in r.''} \emph{Journal of
Statistical Software} 45 (3): 1--67.
\url{https://doi.org/10.18637/jss.v045.i03}.

\leavevmode\vadjust pre{\hypertarget{ref-Cerasonietal2024}{}}%
Cerasoni, J N, M A Bell, and Y E Stuart. 2024. {``Geology,
Microstratigraphy, and Paleontology of Truckee Formation Lacustrine
Diatomite Deposits Near Hazen, Nevada, USA, with Emphasis on Fossil
Stickleback Fish.''} \emph{PaleoBios} 41: 1--15.

\leavevmode\vadjust pre{\hypertarget{ref-Cooperetal2011}{}}%
Cooper, I A, R T Gilman, and J W Boughman. 2011. {``Sexual Dimorphism
and Speciation on Two Ecological Coins: Patterns from Nature and
Theoretical Predictions.''} \emph{Evolution} 65 (9): 2553--71.

\leavevmode\vadjust pre{\hypertarget{ref-DeLisleetal2018}{}}%
De Lisle, S P, S Paiva, and L Rowe. 2018. {``Habitat Partitioning During
Character Displacement Between the Sexes.''} \emph{Biology Letters} 14:
20180124.

\leavevmode\vadjust pre{\hypertarget{ref-DeLisleandRowe2015}{}}%
De Lisle, S P, and L Rowe. 2015. {``Ecological Character Displacement
Between the Sexes.''} \emph{The American Naturalist} 186: 693--707.

\leavevmode\vadjust pre{\hypertarget{ref-DeLisleandRowe2017}{}}%
---------. 2017. {``Disruptive Natural Selection Predicts Divergence
Between the Sexes During Adaptive Radiation.''} \emph{Ecology and
Evolution} 7: 3590--3601.

\leavevmode\vadjust pre{\hypertarget{ref-Godfreyetal1993}{}}%
Godfrey, L R, S K Lyon, and M R Sutherland. 1993. {``Sexual Dimorphism
in Large-Bodied Primates: The Case of the Subfossil Lemurs.''}
\emph{American Journal of Physical Anthropology} 90: 315--34.

\leavevmode\vadjust pre{\hypertarget{ref-Hendryetal2009}{}}%
Hendry, A P, D I Bolnick, D Berner, and C L Peichel. 2009. {``Along the
Speciation Continuum in Sticklebacks.''} \emph{Journal of Fish Biology}
75 (8): 2000--2036.

\leavevmode\vadjust pre{\hypertarget{ref-Herrmannetal2021}{}}%
Herrmann, N C, J T Stroud, and J B Losos. 2021. {``The Evolution of
'Ecological Release' into the 21st Century.''} \emph{Trends in Ecology
and Evolution} 36 (3): 206--15.

\leavevmode\vadjust pre{\hypertarget{ref-HoneandMallon2017}{}}%
Hone, D W E, and J C Mallon. 2017. {``Protracted Growth Impedes the
Detection of Sexual Dimorphism in Non-Avian Dinosaurs.''}
\emph{Palaeontology} 60 (4): 535--45.

\leavevmode\vadjust pre{\hypertarget{ref-Huntetal2008}{}}%
Hunt, G, M A Bell, and M P Travis. 2008. {``Evolution Toward a New
Adaptive Optimum: Phenotypic Evolution in a Fossil Stickleback
Lineage.''} \emph{Evolution} 62 (3): 700--710.

\leavevmode\vadjust pre{\hypertarget{ref-KendallsTau}{}}%
KENDALL, M. G. 1938. {``{A NEW MEASURE OF RANK CORRELATION}.''}
\emph{Biometrika} 30 (1-2): 81--93.
\url{https://doi.org/10.1093/biomet/30.1-2.81}.

\leavevmode\vadjust pre{\hypertarget{ref-KoscinskiandPietraszewski2004}{}}%
Koscinski, K, and S Pietraszewski. 2004. {``Methods to Estimate Sexual
Dimorphism from Unsexed Samples: A Test with Computer Generated
Samples.''} \emph{Przeglad Antropologiczny - Anthropological Review} 67:
33--55.

\leavevmode\vadjust pre{\hypertarget{ref-little2002statistical}{}}%
Little, R. J. A., and D. B. Rubin. 2002. \emph{Statistical Analysis with
Missing Data}. Wiley Series in Probability and Mathematical Statistics.
Probability and Mathematical Statistics. Wiley.
\url{http://books.google.com/books?id=aYPwAAAAMAAJ}.

\leavevmode\vadjust pre{\hypertarget{ref-Mallon2017}{}}%
Mallon, J C. 2017. {``Recognizing Sexual Dimorphism in the Fossil
Record: Lessons from Nonavian Dinosaurs.''} \emph{Paleobiology} 43 (3):
495--507.

\leavevmode\vadjust pre{\hypertarget{ref-ParentandCrespi2009}{}}%
Parent, C E, and B J Crespi. 2009. {``Ecological Opportunity in Adaptive
Radiation of Galapagos Endemic Land Snails.''} \emph{The American
Naturalist} 174: 898--905.

\leavevmode\vadjust pre{\hypertarget{ref-PfennigandPfennig2012}{}}%
Pfennig, D W, and K S Pfennig. 2012. \emph{\href{}{Evolution's Wedge:
Competition and the Origins of Diversity}}. University of California
Press, Berkeley, USA.

\leavevmode\vadjust pre{\hypertarget{ref-Reimchen1994}{}}%
Reimchen, T E. 1994. {``Predators and Morphological Evolution in
Threespine Stickleback.''} In \emph{The Evolutionary Biology of the
Threespine Stickleback}, edited by M A Bell and S A Foster. Oxford, UK:
Oxford University Press.

\leavevmode\vadjust pre{\hypertarget{ref-ReimchenandNosil2004}{}}%
Reimchen, T E, and P Nosil. 2004. {``Variable Predation Regimes Predict
the Evolution of Sexual Dimorphism in a Population of Threespine
Stickleback.''} \emph{Evolution} 58 (6): 1274--81.

\leavevmode\vadjust pre{\hypertarget{ref-Roestietal2023}{}}%
Roesti, M, J S Groh, S A Blain, M Huss, P Rassias, D I Bolnick, Y E
Stuart, C L Peichel, and D Schluter. 2023. {``Species Divergence Under
Competition and Shared Predation.''} \emph{Ecology Letters} 26: 111--23.

\leavevmode\vadjust pre{\hypertarget{ref-Saittaetal2020}{}}%
Saitta, T, M T Stockdale, N R Longrich, V Bonhomme, M J Benton, I C
Cuthill, and P J Makovicky. 2020. {``An Effect Size Statistical
Framework for Investigating Sexual Dimorphism in Non-Avian Dinosaurs and
Other Extinct Taxa.''} \emph{Biological Journal of the Linnean Society}
131 (2): 231--73. \url{https://doi.org/10.1093/biolinnean/blaa105}.

\leavevmode\vadjust pre{\hypertarget{ref-Schindelinetal2012}{}}%
Schindelin, J, I Arganda-Carreras, E Frise, V Kaynig, M Longair, T
Pietzsch, S Preibisch, C Rueden, S Saalfeld, and B Schmid. 2012.
{``Fiji: An Open-Source Platform for Biological-Image Analysis.''}
\emph{Nature Methods} 9: 976--682.

\leavevmode\vadjust pre{\hypertarget{ref-SchluterandMcPhail1992}{}}%
Schluter, D, and J D McPhail. 1992. {``Ecological Character Displacement
and Speciation in Sticklebacks.''} \emph{The American Naturalist} 140
(1): 85--108.

\leavevmode\vadjust pre{\hypertarget{ref-Schoener1968}{}}%
Schoener, T W. 1968. {``The Anolis Lizards of Bimini: Resource
Partitioning in a Complex Fauna.''} \emph{Ecology} 29: 704--26.

\leavevmode\vadjust pre{\hypertarget{ref-Shine1989}{}}%
Shine, R. 1989. {``Ecological Causes for the Evolution of Sexual
Dimorphism: A Review of the Evidence.''} \emph{The Quarterly Review of
Biology} 64 (4): 419--61.

\leavevmode\vadjust pre{\hypertarget{ref-Siddiquietal2024}{}}%
Siddiqui, R, S Swank, A Ozark, F Joaquin, M P Travis, C D McMahan, M A
Bell, and Y E Stuart. 2024. {``Inferring the Evolution of Reproductive
Isolation in a Lineage of Fossil Threespine Stickleback, Gasterosteus
Doryssus.''} \emph{Proceedings of the Royal Society B} 291: 20240337.

\leavevmode\vadjust pre{\hypertarget{ref-Stuartetal2021}{}}%
Stuart, Y E, J W Sherwin, A Kamath, and T Veen. 2021. {``Male and Female
Anolis Carolinensis Maintain Their Dimorphism Despite the Presence of
Novel Interspecific Competition.''} \emph{Evolution} 75 (11): 2708--16.

\leavevmode\vadjust pre{\hypertarget{ref-Stuartetal2020}{}}%
Stuart, Y E, M P Travis, and M A Bell. 2020. {``Inferred Genetic
Architecture Underlying Evolution in a Fossil Stickleback Lineage.''}
\emph{Nature Ecology and Evolution} 4 (11): 1549--57.

\leavevmode\vadjust pre{\hypertarget{ref-Stuartetal2017}{}}%
Stuart, Y E, T Veen, J N Weber, D Hanson, M Ravinet, B K Lohman, C J
Thompson, et al. 2017. {``Contrasting Effects of Environment and
Genetics Generate a Continuum of Parallel Evolution.''} \emph{Nature
Ecology and Evolution} 1 (6): 158.

\leavevmode\vadjust pre{\hypertarget{ref-R2022language}{}}%
Team, R Core. 2022. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\leavevmode\vadjust pre{\hypertarget{ref-OUProcess}{}}%
Uhlenbeck, G E, and L S Ornstein. 1930. {``On the Theory of the Brownian
Motion.''} \emph{Phys. Rev.} 36 (September): 823--41.
\url{https://doi.org/10.1103/PhysRev.36.823}.

\leavevmode\vadjust pre{\hypertarget{ref-VamosiandSchluter2004}{}}%
Vamosi, S M, and D Schluter. 2004. {``Character Shifts in the Defensive
Armor of Sympatric Sticklebacks.''} \emph{Evolution} 58: 376--85.

\leavevmode\vadjust pre{\hypertarget{ref-Vojeetal2022}{}}%
Voje, K L, M A Bell, and Y E Stuart. 2022. {``Evolution of Static
Allometry and Constraint on Evolutionary Allometry in a Fossil
Stickleback.''} \emph{Journal of Evolutionary Biology} 35 (3): 423--38.

\leavevmode\vadjust pre{\hypertarget{ref-ZhouReiter2010}{}}%
Zhou, X, and J Reiter. 2010. {``A Note on Bayesian Inference After
Multiple Imputation.''} \emph{The American Statistician} 64 (2):
159--63.

\end{CSLReferences}

\end{document}